---
title: "Linking Nielsen Scanner Data and FDC Branded Foods"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## LINKING APPROACH: use last 12 digits of gtin_upc

# note: this approach of mine did not go very well...

# Setup

```{r libraries}
library(readr)
library(dplyr)
library(stringr)
library(duckdb)
library(DBI)
```

```{r paths}
nielsen_root <- "G:/Shared drives/Nielsen Scanner/Data"
path_stores      <- file.path(nielsen_root, "01_Annual_Files/stores.tsv")
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")
path_vegetables  <- file.path(nielsen_root, "06_Groceries/VEGETABLES.tsv")

lab_dir <- "c:/Users/hughc/OneDrive/Desktop/Spring 2026/FAE Lab"
fdc_branded_csv <- file.path(
  lab_dir,
  "FoodData_Central_branded_food_csv_2026-04-30",
  "branded_food.csv"
)

analysis_week <- "2023-05-13"
n_upcs <- 1000L

path_nielsen_upcs <- file.path(lab_dir, "nielsen_vegetables_upcs.csv")
path_linked       <- file.path(lab_dir, "nielsen_fdc_linked.csv")
```

```{r helpers}
normalize_upc12 <- function(x) {
  digits <- str_remove_all(as.character(x), "[^0-9]")
  digits <- ifelse(is.na(digits) | digits == "", NA_character_, digits)
  str_sub(str_pad(digits, width = 12, side = "left", pad = "0"), -12L)
}
```

# Nielsen vegetable UPCs

Massachusetts stores, one analysis week, top `r n_upcs` UPCs by total units sold.

```{r nielsen-upcs}
stores <- read_tsv(path_stores, show_col_types = FALSE)
ma_store_codes <- stores %>%
  filter(fips_state_code == 25) %>%
  pull(store_code_uc)

con <- dbConnect(duckdb())
on.exit(dbDisconnect(con, shutdown = TRUE), add = TRUE)

dbWriteTable(con, "ma_store_codes", data.frame(store_code_uc = ma_store_codes), temporary = TRUE)

nielsen_upcs <- dbGetQuery(con, sprintf("
  WITH veg AS (
    SELECT
      v.upc,
      p.product_descr,
      COUNT(*) AS n_store_rows,
      SUM(v.units) AS total_units
    FROM read_csv_auto('%s', delim = '\\t') v
    LEFT JOIN read_csv_auto('%s', delim = '\\t') p
      ON v.upc = p.upc
    WHERE v.store_code_uc IN (SELECT store_code_uc FROM ma_store_codes)
      AND v.week_end = '%s'
    GROUP BY v.upc, p.product_descr
  ),
  by_upc AS (
    SELECT
      upc,
      ANY_VALUE(product_descr) AS product_descr,
      SUM(n_store_rows) AS n_store_rows,
      SUM(total_units) AS total_units
    FROM veg
    GROUP BY upc
  )
  SELECT *
  FROM by_upc
  ORDER BY total_units DESC
  LIMIT %d
", path_vegetables, path_productdesc, analysis_week, n_upcs)) %>%
  mutate(upc_norm = normalize_upc12(upc))

write_csv(nielsen_upcs, path_nielsen_upcs)
head(nielsen_upcs)
```

# FDC Branded Foods lookup

```{r fdc-branded}
dbExecute(con, sprintf("
  CREATE OR REPLACE TABLE fdc_branded AS
  SELECT
    fdc_id,
    brand_owner,
    gtin_upc,
    branded_food_category,
    ingredients,
    serving_size,
    serving_size_unit,
    modified_date,
    available_date,
    discontinued_date,
    right(
      lpad(regexp_replace(cast(gtin_upc AS VARCHAR), '[^0-9]', '', 'g'), 12, '0'),
      12
    ) AS upc_norm
  FROM read_csv_auto('%s')
  WHERE gtin_upc IS NOT NULL
    AND trim(cast(gtin_upc AS VARCHAR)) <> ''
", fdc_branded_csv))
```

# Merge

```{r merge}
dbWriteTable(con, "nielsen_upcs", nielsen_upcs, overwrite = TRUE, temporary = TRUE)

nielsen_fdc_linked <- dbGetQuery(con, "
  SELECT
    n.upc AS nielsen_upc,
    n.product_descr AS nielsen_product_descr,
    n.total_units,
    n.n_store_rows,
    f.fdc_id,
    f.gtin_upc,
    f.brand_owner,
    f.branded_food_category,
    f.ingredients,
    f.modified_date,
    f.discontinued_date
  FROM nielsen_upcs n
  LEFT JOIN fdc_branded f
    ON n.upc_norm = f.upc_norm
  ORDER BY n.upc, f.fdc_id
")

write_csv(nielsen_fdc_linked, path_linked)

link_summary <- tibble(
  nielsen_upcs = nrow(nielsen_upcs),
  upcs_matched = n_distinct(nielsen_fdc_linked$nielsen_upc[!is.na(nielsen_fdc_linked$fdc_id)]),
  total_link_rows = sum(!is.na(nielsen_fdc_linked$fdc_id)),
  upcs_multiple_fdc = nielsen_fdc_linked %>%
    filter(!is.na(fdc_id)) %>%
    count(nielsen_upc) %>%
    filter(n > 1) %>%
    nrow()
) %>%
  mutate(pct_matched = round(100 * upcs_matched / nielsen_upcs, 1))

link_summary
```

One Nielsen UPC can match multiple FDC records (package sizes, reformulations). For inspection, keep the first match per UPC:

```{r one-per-upc}
nielsen_fdc_one <- nielsen_fdc_linked %>%
  group_by(nielsen_upc) %>%
  slice(1) %>%
  ungroup()

head(
  nielsen_fdc_one %>%
    select(
      nielsen_upc, nielsen_product_descr,
      gtin_upc, brand_owner, branded_food_category
    )
)
```

# Unmatched UPCs

```{r unmatched}
nielsen_fdc_one %>%
  filter(is.na(fdc_id)) %>%
  arrange(desc(total_units)) %>%
  select(nielsen_upc, nielsen_product_descr, total_units) %>%
  head(15)
```
