---
title: "Testing Code - 2026"
output: html_document
date: "2026-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## README: For steps 1 and 3 I am unsure if we want to work with a big merged dataframe or a list of dataframes, so I kind of do both

# Load required libraries
```{r }
library(installr) # Making sure we have latest R version

library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stringr)

library(duckdb)
library(DBI)

library(purrr)
library(tools)

# updateR()
```

## Load in Nielsen Data
```{r}
stores <- read_tsv("G:\\Shared drives\\Nielsen Scanner\\Data\\01_Annual_Files\\stores.tsv")
product_attri_prev <- read_tsv("G:\\Shared drives\\Nielsen Scanner\\Data\\01_Annual_Files\\productattributes.tsv", n_max = 100)

```



## Step 1: Load in the two week MA data and merge
```{r}
load_tag_and_merge <- function(folder, files = NULL, ...) {
  if (is.null(files)) {
    files <- list.files(folder, pattern = "\\.csv$", full.names = FALSE)
  }
  
  # get a list of paths
  paths <- file.path(folder, files)

  # add column to each row specifying what csv it came from
  dfs <- lapply(paths, function(p) {
    df <- read.csv(p, ...)
    df$type <- sub("_.*", "", tools::file_path_sans_ext(basename(p)))
    df
  })
  names(dfs) <- tools::file_path_sans_ext(basename(files))

  # merge all the datasets
  merged <- do.call(rbind, dfs)
  rownames(merged) <- NULL

  list(dfs = dfs, merged = merged)
}

MA_twoweeks <- "G:\\Shared drives\\Nielsen Scanner\\Data\\MA_twoweeks"

result <- load_tag_and_merge(MA_twoweeks)
dfs    <- result$dfs      # list of data.frames (one per category CSV)
merged <- result$merged   # one combined data.frame (all stores, all categories)

# NOTE: upc's are stored as numeric so they may drop the leading 0's
merged <- merged %>%
  mutate(upc_chr = sprintf("%.0f", upc))

# Making one data frame per store
by_store <- split(merged, merged$store_code_uc)

length(by_store)                 # number of MA stores in this extract
names(by_store)[1:5]             # store codes as list names
head(by_store[[1]])              # peek at one store
```

Keep `merged` as the full panel. Split into a **named list** keyed by `store_code_uc` for per-store work (HEI, variety scores, etc.).


## Step 2: Link Nielsen UPCs to Jus FNDDS/branded data (`newdf`) — diagnostics + multi-rule match

```{r step2-setup}
lab_dir <- "c:/Users/hughc/OneDrive/Desktop/Spring 2026/FAE Lab"
nielsen_root <- "G:/Shared drives/Nielsen Scanner/Data"
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

# --- helpers: digit strings and join keys ---
digits_only <- function(x) {
  d <- str_remove_all(as.character(x), "[^0-9]")
  ifelse(is.na(d) | d == "", NA_character_, d)
}

pad12_last <- function(d) {
  str_sub(str_pad(d, 12, side = "left", pad = "0"), -12L)
}

add_nielsen_keys <- function(df, upc_col = "upc_chr") {
  d <- digits_only(df[[upc_col]])
  df %>%
    mutate(
      upc_digits = d,
      upc_digit_len = nchar(upc_digits),
      key_last12 = pad12_last(upc_digits)
    )
}

add_gtin_keys <- function(df, gtin_col = "gtin_upc") {
  d <- digits_only(df[[gtin_col]])
  df %>%
    mutate(
      gtin_digits = d,
      gtin_digit_len = nchar(gtin_digits),
      key_last12 = pad12_last(gtin_digits),
      key_gtin13_body = if_else(gtin_digit_len == 13L, str_sub(gtin_digits, 1, 12), NA_character_),
      key_gtin14_body = if_else(gtin_digit_len == 14L, str_sub(gtin_digits, 2, 13), NA_character_),
      # Force character even if column is all NA (prevents logical type)
      key_exact = as.character(gtin_digits)
    )
}
```

```{r step2-foodgroups, message = FALSE}
# Jus pipeline -> object `newdf` (FOODCODE, nutrients, MyPlate flags, fdc_id, gtin_upc)
setwd(lab_dir)
source("foodgroups.R")
```

```{r step2-gtin-audit}
# --- A) GTIN quality in Jus `newdf`
gtin_audit_newdf <- newdf %>%
  mutate(
    gtin_digits = digits_only(gtin_upc),
    gtin_digit_len = nchar(gtin_digits)
  ) %>%
  summarise(
    n_fndds_rows = n(),
    n_missing_gtin = sum(is.na(gtin_upc) | str_trim(gtin_upc) == ""),
    n_with_gtin = sum(!is.na(gtin_upc) & str_trim(gtin_upc) != ""),
    n_with_digits = sum(!is.na(gtin_digits)),
    pct_missing_gtin = round(100 * n_missing_gtin / n_fndds_rows, 1),
    n_unique_gtin = n_distinct(gtin_digits, na.rm = TRUE),
    n_unique_fdc_id = n_distinct(fdc_id, na.rm = TRUE)
  )

gtin_len_dist_newdf <- newdf %>%
  mutate(gtin_digit_len = nchar(digits_only(gtin_upc))) %>%
  count(gtin_digit_len, name = "n_rows", sort = TRUE) %>%
  mutate(pct = round(100 * n_rows / sum(n_rows), 1))

# Duplicate GTINs in newdf (many FNDDS codes can share one barcode)
gtin_dupes_newdf <- newdf %>%
  mutate(gtin_digits = digits_only(gtin_upc)) %>%
  filter(!is.na(gtin_digits)) %>%
  count(gtin_digits, name = "n_fndds_rows") %>%
  filter(n_fndds_rows > 1) %>%
  arrange(desc(n_fndds_rows))

print(gtin_audit_newdf)
print(head(gtin_len_dist_newdf, 10))
cat("GTINs linked to >1 FNDDS row:", nrow(gtin_dupes_newdf), "\n")
```

```{r step2-nielsen-products}
# --- B) Nielsen product catalog from our bi-weekly `merged` panel ---
nielsen_products <- merged %>%
  group_by(upc_chr, type) %>%
  summarise(
    n_stores = n_distinct(store_code_uc),
    total_units_sold = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  add_nielsen_keys()

# Optional: attach Nielsen product descriptions (for "NO PACKAGE" diagnostics)
if (file.exists(path_productdesc)) {
  con_desc <- dbConnect(duckdb())

  dbWriteTable(
    con_desc, "nielsen_upc_list",
    nielsen_products %>%
      distinct(upc_digits) %>%
      filter(!is.na(upc_digits), upc_digits != ""),
    temporary = TRUE
  )

  # Match on digits only (productdesc upc may be string or numeric in the TSV)
  path_productdesc_sql <- gsub("'", "''", path_productdesc)
  product_desc <- dbGetQuery(con_desc, sprintf("
    SELECT
      regexp_replace(trim(cast(p.upc AS VARCHAR)), '[^0-9]', '', 'g') AS upc_digits,
      any_value(p.product_descr) AS product_descr
    FROM read_csv_auto('%s', delim = '\\t', header = TRUE) p
    INNER JOIN nielsen_upc_list u
      ON regexp_replace(trim(cast(p.upc AS VARCHAR)), '[^0-9]', '', 'g') = u.upc_digits
    GROUP BY 1
  ", path_productdesc_sql))

  dbDisconnect(con_desc, shutdown = TRUE)

  nielsen_products <- nielsen_products %>%
    left_join(product_desc, by = "upc_digits") %>%
    mutate(
      is_no_package = str_detect(toupper(coalesce(product_descr, "")), "NO PACKAGE")
    )
} else {
  nielsen_products <- nielsen_products %>%
    mutate(product_descr = NA_character_, is_no_package = NA)
  warning("productdesc.tsv not found; is_no_package will be NA.")
}
```

```{r step2-fdc-lookup}
# --- C) FDC / FNDDS lookup from `newdf` (one row per gtin_digits + fdc_id) ---
fdc_cols <- c(
  "fdc_id", "gtin_upc", "FOODCODE", "DESCRIPTION",
  "KCAL", "SODIUM", "SATFAT", "MUFA", "PUFA",
  "is_veg", "is_fruit", "is_whole_grain", "is_refined_grain",
  "is_protein", "is_dairy", "is_oil", "is_added_sugar", "is_solid_fat"
)
fdc_cols <- intersect(fdc_cols, names(newdf))

fdc_lookup <- newdf %>%
  select(all_of(fdc_cols)) %>%
  filter(!is.na(gtin_upc), str_trim(gtin_upc) != "") %>%
  add_gtin_keys() %>%
  mutate(
    across(
      any_of(c("key_last12", "key_gtin13_body", "key_gtin14_body", "key_exact")),
      as.character
    )
  ) %>%
  distinct(fdc_id, gtin_digits, .keep_all = TRUE)
```

```{r step2-match-rules}
# --- D) Try several join rules; record which rule hit ---
match_rules <- list(
  last12_eq_last12       = c("key_last12", "key_last12"),
  nielsen12_eq_gtin13body = c("key_last12", "key_gtin13_body"),
  nielsen12_eq_gtin14body = c("key_last12", "key_gtin14_body"),
  exact_digits           = c("upc_digits", "key_exact")
)

match_hits <- imap_dfr(match_rules, function(cols, rule_name) {
  n_key <- cols[1]
  f_key <- cols[2]

  nielsen_products %>%
    filter(!is.na(.data[[n_key]])) %>%
    inner_join(
      fdc_lookup %>% filter(!is.na(.data[[f_key]])),
      by = setNames(f_key, n_key),
      relationship = "many-to-many"
    ) %>%
    transmute(
      upc_chr, type, match_rule = rule_name,
      n_stores, total_units_sold,
      product_descr, is_no_package,
      fdc_id, gtin_upc, FOODCODE, DESCRIPTION,
      across(any_of(c(
        "is_veg", "is_fruit", "is_whole_grain", "is_refined_grain",
        "is_protein", "is_dairy", "is_oil", "is_added_sugar", "is_solid_fat",
        "KCAL", "SODIUM", "SATFAT", "MUFA", "PUFA"
      )))
    )
})

rule_priority <- c(
  "exact_digits",
  "nielsen12_eq_gtin13body",
  "last12_eq_last12",
  "nielsen12_eq_gtin14body"
)

# One best row per Nielsen product (upc_chr + type)
nielsen_fdc_best <- match_hits %>%
  mutate(match_rule = factor(match_rule, levels = rule_priority)) %>%
  group_by(upc_chr, type) %>%
  arrange(match_rule, desc(total_units_sold), FOODCODE) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(matched = TRUE)

# --- E) Match-rate summary: overall, by rule, by category, by NO PACKAGE ---
n_distinct_upc <- nielsen_products %>% distinct(upc_chr) %>% nrow()

match_summary_overall <- tibble(
  n_nielsen_upc_type = nrow(nielsen_products),
  n_distinct_upc = n_distinct_upc,
  n_matched_upc_type = nrow(nielsen_fdc_best),
  n_matched_distinct_upc = n_distinct(nielsen_fdc_best$upc_chr),
  pct_matched_upc_type = round(100 * n_matched_upc_type / n_nielsen_upc_type, 1),
  pct_matched_distinct_upc = round(100 * n_matched_distinct_upc / n_distinct_upc, 1)
)

match_by_rule <- match_hits %>%
  distinct(upc_chr, type, match_rule) %>%
  count(match_rule, name = "n_upc_type") %>%
  mutate(pct = round(100 * n_upc_type / nrow(nielsen_products), 1)) %>%
  arrange(desc(n_upc_type))

match_by_type <- nielsen_fdc_best %>%
  count(type, name = "n_matched") %>%
  right_join(count(nielsen_products, type), by = "type") %>%
  rename(n_total = n) %>%
  mutate(
    n_matched = coalesce(n_matched, 0L),
    pct_matched = round(100 * n_matched / n_total, 1)
  ) %>%
  arrange(type)

match_by_no_package <- if (all(is.na(nielsen_products$is_no_package))) {
  tibble(note = "product_descr not available")
} else {
  nielsen_products %>%
    left_join(
      nielsen_fdc_best %>% select(upc_chr, type, fdc_id),
      by = c("upc_chr", "type")
    ) %>%
    mutate(matched = !is.na(fdc_id)) %>%
    group_by(is_no_package) %>%
    summarise(
      n_products = n(),
      n_matched = sum(matched),
      pct_matched = round(100 * n_matched / n_products, 1),
      total_units = sum(total_units_sold),
      .groups = "drop"
    )
}

match_summary_overall
match_by_rule
match_by_type
match_by_no_package
```

```{r step2-merge-to-panel}
# --- F) Attach best match to transaction-level `merged` and product summary ---
merged_linked <- merged %>%
  left_join(
    nielsen_fdc_best %>%
      select(
        upc_chr, type, match_rule, fdc_id, gtin_upc, FOODCODE, DESCRIPTION,
        starts_with("is_"), KCAL, SODIUM, SATFAT, MUFA, PUFA
      ),
    by = c("upc_chr", "type")
  ) %>%
  mutate(matched_fdc = !is.na(fdc_id))

# For Step 3: product_summary with match flags
product_summary <- nielsen_products %>%
  left_join(
    nielsen_fdc_best %>%
      select(upc_chr, type, match_rule, fdc_id, FOODCODE, matched),
  by = c("upc_chr", "type")
  ) %>%
  mutate(matched = coalesce(matched, FALSE))

# Save diagnostics (optional)
write_csv(match_summary_overall, file.path(lab_dir, "step2_match_summary_overall.csv"))
write_csv(match_by_rule, file.path(lab_dir, "step2_match_by_rule.csv"))
write_csv(match_by_type, file.path(lab_dir, "step2_match_by_type.csv"))
write_csv(
  nielsen_fdc_best %>%
    select(upc_chr, type, match_rule, fdc_id, gtin_upc, FOODCODE, DESCRIPTION, product_descr),
  file.path(lab_dir, "step2_nielsen_fdc_best.csv")
)

cat("Rows in merged_linked:", nrow(merged_linked),
    "| matched rows:", sum(merged_linked$matched_fdc),
    "(", round(100 * mean(merged_linked$matched_fdc), 1), "%)\n")
```

## Step 3: product-level summary (one merged table)

`product_summary` is built in Step 2 (store counts, units sold, FDC match flags). Re-run Step 1 alone if you only need the base summary without FDC columns.

```{r step3-product-summary}
product_summary <- product_summary %>%
  arrange(type, desc(n_stores), desc(total_units_sold))

head(product_summary, 20)
```

Later, top-20-by-type is easier from a **list** (one small table per `type`) than from one giant ranked table:

```{r step3-top20-by-type-preview}
top20_by_type <- product_summary %>%
  group_by(type) %>%
  slice_max(order_by = total_units_sold, n = 20, with_ties = FALSE) %>%
  ungroup()

# Or split into a list: top20_by_type_list[["VEGETABLES"]], etc.
top20_by_type_list <- split(
  top20_by_type,
  top20_by_type$type
)
```

