---
title: "Testing Code - 2026"
output:
  pdf_document: default
date: "2026-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r paths}
# Change this to your working directory
lab_dir <- "C:/Users/hughc/Spring 2026/FAE Lab"

nielsen_root <- "G:/Shared drives/Nielsen Scanner/Data"
path_stores <- file.path(nielsen_root, "01_Annual_Files/stores.tsv")
path_ma_twoweeks <- file.path(nielsen_root, "MA_twoweeks")
path_branded_food <- file.path(lab_dir, "dataraw/branded_food.csv")
path_food_desc    <- file.path(lab_dir, "FoodData_Central_branded_food_csv_2026-04-30/food.csv")

week_baseline <- as.Date("2023-05-13")   # May — primary outputs
week_alt      <- as.Date("2023-09-30")   # September — parallel outputs
```

```{r libraries}
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(tools)
library(readxl)
library(purrr)
```

```{r helpers}
digits_only <- function(x) {
  d <- str_remove_all(as.character(x), "[^0-9]")
  ifelse(is.na(d) | d == "", NA_character_, d)
}

pad12_last <- function(d) {
  str_sub(str_pad(d, 12, side = "left", pad = "0"), -12L)
}

add_nielsen_keys <- function(df, upc_col = "upc_chr") {
  d <- digits_only(df[[upc_col]])
  df %>%
    mutate(
      upc_digits = d,
      key_last12 = pad12_last(upc_digits),
      upc_trunc = str_sub(key_last12, 3, -1)
    )
}

# Returns total basket cost (and item count) broken out by period.
# Input: one store's basket dataframe (or any slice of store_basket).
# Each row = one product for one week, so sum(price) per period = cost of
# buying one unit of every item in the basket during that sampling week.
basket_total_price <- function(basket_df) {
  basket_df %>%
    group_by(week_end) %>%
    summarise(
      n_items     = n(),
      total_price = sum(price, na.rm = TRUE),
      .groups     = "drop"
    )
}

# Branded Foods serving_size is usually g or ml; standardize for nutrient/HEI prep
add_serving_units <- function(df) {
  df %>%
    mutate(
      serving_size_num = as.numeric(serving_size),
      serving_unit_raw = tolower(str_trim(serving_size_unit)),
      serving_size_g = case_when(
        serving_unit_raw %in% c("g", "grm", "gram", "grams") ~ serving_size_num,
        serving_unit_raw %in% c("oz", "onz") ~ serving_size_num * 28.3495,
        TRUE ~ NA_real_
      ),
      serving_size_ml = if_else(
        serving_unit_raw %in% c("ml", "mlt"),
        serving_size_num,
        NA_real_
      ),
      serving_basis = case_when(
        serving_unit_raw %in% c("ml", "mlt") ~ "ml",
        !is.na(serving_size_g) ~ "g",
        TRUE ~ NA_character_
      )
    )
}
```

# Step 1: Nielsen two-week panel (MA)

```{r stores}
stores <- read_tsv(path_stores, show_col_types = FALSE)
```

```{r step1-load}
load_tag_and_merge <- function(folder, files = NULL, ...) {
  if (is.null(files)) {
    files <- list.files(folder, pattern = "\\.csv$", full.names = FALSE)
  }
  paths <- file.path(folder, files)
  dfs <- lapply(paths, function(p) {
    df <- read.csv(p, ...)
    df$type <- sub("_.*", "", tools::file_path_sans_ext(basename(p)))
    df
  })
  names(dfs) <- tools::file_path_sans_ext(basename(files))
  merged <- do.call(rbind, dfs)
  rownames(merged) <- NULL
  list(dfs = dfs, merged = merged)
}

result <- load_tag_and_merge(path_ma_twoweeks)
dfs <- result$dfs
merged <- result$merged %>%
  mutate(upc_chr = sprintf("%.0f", upc))

by_store <- split(merged, merged$store_code_uc)
length(by_store)
head(merged)
```

# Step 2: Link Nielsen to Branded Foods

UPC match: Nielsen `str_sub(key_last12, 3, -1)` = Branded `str_sub(gtin_upc, 2, -2)` for 12-character GTINs.

FNDDS/FPED are not joined here; `fdc_id` is retained for a future link to survey nutrients.

```{r step2-branded}
branded_food <- read_csv(path_branded_food, col_types = cols(.default = "c"))

branded_for_join <- branded_food %>%
  mutate(gtin_upc_chr = as.character(gtin_upc)) %>%
  filter(nchar(gtin_upc_chr) == 12) %>%
  mutate(gtin_upc_trunc = str_sub(gtin_upc_chr, 2, -2)) %>%
  select(
    fdc_id,
    gtin_upc = gtin_upc_chr,
    gtin_upc_trunc,
    brand_owner,
    brand_name,
    branded_food_category,
    short_description,
    serving_size,
    serving_size_unit,
    household_serving_fulltext
  )
```

```{r step2-link}
nielsen_products <- merged %>%
  group_by(upc_chr, type) %>%
  summarise(
    n_stores = n_distinct(store_code_uc),
    total_units_sold = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  add_nielsen_keys()

nielsen_fdc_linked <- nielsen_products %>%
  select(upc_chr, type, n_stores, total_units_sold, upc_trunc) %>%
  full_join(
    branded_for_join,
    by = join_by(upc_trunc == gtin_upc_trunc),
    relationship = "many-to-many"
  )

nielsen_fdc_best <- nielsen_fdc_linked %>%
  filter(!is.na(fdc_id)) %>%
  group_by(upc_chr, type) %>%
  slice(1) %>%
  ungroup()

branded_cols <- c(
  "fdc_id", "gtin_upc", "brand_owner", "brand_name",
  "branded_food_category", "short_description",
  "serving_size", "serving_size_unit", "household_serving_fulltext"
)

product_catalog <- nielsen_fdc_best %>%
  select(upc_chr, type, all_of(branded_cols)) %>%
  add_serving_units()

merged_linked <- merged %>%
  left_join(product_catalog, by = c("upc_chr", "type")) %>%
  mutate(matched_fdc = !is.na(fdc_id))

product_summary <- nielsen_products %>%
  left_join(
    nielsen_fdc_best %>% select(upc_chr, type, all_of(branded_cols)),
    by = c("upc_chr", "type")
  ) %>%
  mutate(matched = !is.na(fdc_id))

tibble(
  n_nielsen_products = nrow(nielsen_products),
  n_matched = nrow(nielsen_fdc_best),
  pct_matched = round(100 * nrow(nielsen_fdc_best) / nrow(nielsen_products), 1),
  n_merged_rows = nrow(merged_linked),
  n_merged_matched = sum(merged_linked$matched_fdc)
)
```

# Step 3: Product basket (matched, Food + Convenience)

```{r step3-channels}
channel_labels <- c(
  C = "Convenience",
  F = "Food",
  G = "Convenience with Gas Station",
  D = "Drug",
  M = "Mass Merchandiser",
  L = "Liquor"
)

stores_channels <- stores %>%
  filter(fips_state_code == 25, channel_code %in% c("F", "C")) %>%
  mutate(channel_name = recode(channel_code, !!!channel_labels)) %>%
  select(store_code_uc, channel_code, channel_name)
```

```{r step3-top50}
matched_panel <- merged_linked %>%
  filter(matched_fdc) %>%
  inner_join(stores_channels, by = "store_code_uc")

top50_by_category <- matched_panel %>%
  group_by(type, upc_chr) %>%
  summarise(
    n_stores_carrying = n_distinct(store_code_uc),
    total_units_fc = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(type) %>%
  slice_max(order_by = n_stores_carrying, n = 50, with_ties = FALSE) %>%
  ungroup() %>%
  left_join(product_catalog, by = c("upc_chr", "type"))

top50_by_category %>% count(type, name = "n_products")

write.csv(
  as.data.frame(top50_by_category),
  file.path(lab_dir, "top50_by_category.csv"),
  row.names = FALSE
)
```

```{r step3-store-basket}
store_sales_top50 <- matched_panel %>%
  inner_join(top50_by_category %>% select(type, upc_chr), by = c("type", "upc_chr")) %>%
  group_by(store_code_uc, channel_code, channel_name, type, upc_chr) %>%
  summarise(
    store_units = sum(units, na.rm = TRUE),
    .groups = "drop"
  )

top5_per_store <- store_sales_top50 %>%
  group_by(store_code_uc, channel_code, channel_name, type) %>%
  slice_max(order_by = store_units, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  select(store_code_uc, type, upc_chr)

# One row per store x product x week, carrying Nielsen's own price/units for
# that store-week directly (no cross-store or cross-week averaging).
store_basket <- matched_panel %>%
  inner_join(top5_per_store, by = c("store_code_uc", "type", "upc_chr")) %>%
  mutate(period = format(as.Date(week_end), "%b %Y")) %>%
  select(
    store_code_uc, channel_code, channel_name, type, upc_chr,
    week_end, period, units, price,
    all_of(branded_cols), serving_size_g, serving_size_ml, serving_basis
  ) %>%
  as.data.frame()

store_basket_may <- store_basket %>% filter(as.Date(week_end) == week_baseline)
store_basket_sep <- store_basket %>% filter(as.Date(week_end) == week_alt)

write_baskets <- function(sb, dir_name) {
  write.csv(sb, file.path(lab_dir, paste0(dir_name, ".csv")), row.names = FALSE)
  sb_slim <- sb %>%
    select(-channel_name, -period, -brand_name, -short_description,
           -serving_size, -serving_size_unit, -household_serving_fulltext,
           -serving_size_g, -serving_size_ml, -serving_basis)
  by_store <- lapply(split(sb_slim, sb_slim$store_code_uc), as.data.frame)
  basket_dir <- file.path(lab_dir, dir_name)
  dir.create(basket_dir, showWarnings = FALSE, recursive = TRUE)
  for (sid in names(by_store)) {
    write.csv(by_store[[sid]], file.path(basket_dir, paste0(sid, ".csv")), row.names = FALSE)
  }
  by_store
}

basket_by_store     <- write_baskets(store_basket_may, "baskets_by_store")
basket_by_store_sep <- write_baskets(store_basket_sep, "baskets_by_store_sep2023")

tibble(
  n_top50_rows    = nrow(top50_by_category),
  n_basket_may    = nrow(store_basket_may),
  n_basket_sep    = nrow(store_basket_sep),
  n_stores        = length(basket_by_store)
)
```

```{r step3-prices}
# Median price per product per period across all MA Food + Convenience stores.
# Two columns (one per sample week) so the HEI script can look up a reference
# price without needing a specific store.
product_prices <- matched_panel %>%
  inner_join(top50_by_category %>% select(type, upc_chr), by = c("type", "upc_chr")) %>%
  mutate(period_key = paste0("price_", format(as.Date(week_end), "%b%Y"))) %>%
  group_by(type, upc_chr, period_key) %>%
  summarise(price_ref = median(price, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = period_key, values_from = price_ref)

top50_by_category <- top50_by_category %>%
  left_join(product_prices, by = c("type", "upc_chr"))

write.csv(
  as.data.frame(top50_by_category),
  file.path(lab_dir, "top50_by_category.csv"),
  row.names = FALSE
)

# Quick check: how many products have a price for each period?
top50_by_category %>%
  summarise(
    n_products       = n(),
    n_price_May2023  = sum(!is.na(price_May2023)),
    n_price_Sep2023  = sum(!is.na(price_Sep2023))
  )
```

# Week deliverables

## TODO 0 — Matched product list (~500) for FPED work

The lab “~500 products” list is **`top50_by_category`**: matched Nielsen × Branded items, up to **50 per category** (15 categories → **747 rows**; EGGS has 47).

```{r todo0-product-list}
product_list_matched <- top50_by_category

todo0_check <- tibble(
  n_rows_product_list = nrow(product_list_matched),
  n_categories = n_distinct(product_list_matched$type),
  all_have_fdc_id = all(!is.na(product_list_matched$fdc_id)),
  all_have_gtin = all(!is.na(product_list_matched$gtin_upc))
)

todo0_check

write.csv(
  as.data.frame(product_list_matched),
  file.path(lab_dir, "product_list_matched.csv"),
  row.names = FALSE
)

# Full matched catalog (all UPC × type with a branded hit, not only top 50)
write.csv(
  as.data.frame(nielsen_fdc_best),
  file.path(lab_dir, "matched_products_all.csv"),
  row.names = FALSE
)
```

## TODO 1 — Sanity check: unmatched products

```{r todo1-unmatched}
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

unmatched_products <- product_summary %>%
  filter(!matched) %>%
  arrange(type, desc(total_units_sold))

if (file.exists(path_productdesc)) {
  product_desc <- read_tsv(path_productdesc, show_col_types = FALSE) %>%
    mutate(upc_chr = sprintf("%.0f", as.numeric(upc))) %>%
    distinct(upc_chr, .keep_all = TRUE) %>%
    select(upc_chr, product_descr)

  unmatched_products <- unmatched_products %>%
    left_join(product_desc, by = "upc_chr")
}

unmatched_by_type <- unmatched_products %>%
  group_by(type) %>%
  summarise(
    n_unmatched = n(),
    total_units_unmatched = sum(total_units_sold, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(product_summary %>% count(type, name = "n_total"), by = "type") %>%
  mutate(pct_unmatched = round(100 * n_unmatched / n_total, 1)) %>%
  arrange(desc(n_unmatched))

match_rate_by_type <- product_summary %>%
  group_by(type) %>%
  summarise(
    n_total = n(),
    n_matched = sum(matched),
    n_unmatched = sum(!matched),
    pct_matched = round(100 * mean(matched), 1),
    .groups = "drop"
  )

unmatched_by_type
match_rate_by_type

write.csv(as.data.frame(unmatched_products), file.path(lab_dir, "unmatched_products.csv"), row.names = FALSE)
write.csv(as.data.frame(unmatched_by_type), file.path(lab_dir, "unmatched_by_type.csv"), row.names = FALSE)
write.csv(as.data.frame(match_rate_by_type), file.path(lab_dir, "match_rate_by_type.csv"), row.names = FALSE)
```

## TODO 3 — Match sanity check: Nielsen vs Branded Foods descriptions

```{r todo3-match-sanity}
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

if (file.exists(path_productdesc)) {
  product_desc_all <- read_tsv(path_productdesc, show_col_types = FALSE) %>%
    mutate(upc_chr = sprintf("%.0f", as.numeric(upc))) %>%
    distinct(upc_chr, .keep_all = TRUE) %>%
    select(upc_chr, nielsen_desc = product_descr)

  match_check <- top50_by_category %>%
    left_join(product_desc_all, by = "upc_chr") %>%
    mutate(
      branded_label = paste(
        coalesce(brand_owner, ""),
        coalesce(brand_name, ""),
        coalesce(branded_food_category, ""),
        sep = " | "
      )
    ) %>%
    select(
      type,
      upc_chr,
      gtin_upc,
      nielsen_desc,
      brand_owner,
      brand_name,
      branded_label,
      branded_category = branded_food_category
    ) %>%
    arrange(type, upc_chr)

  print(match_check, n = Inf)

  write.csv(
    as.data.frame(match_check),
    file.path(lab_dir, "match_sanity_check.csv"),
    row.names = FALSE
  )
} else {
  message("productdesc.tsv not found — skipping match sanity check")
}
```

## TODO 2 — Serving units (product catalog) + store-level pricing


```{r todo2-product-catalog}
write.csv(as.data.frame(product_catalog), file.path(lab_dir, "product_catalog.csv"), row.names = FALSE)

product_catalog %>%
  summarise(
    n_products = n(),
    n_with_serving_g = sum(!is.na(serving_size_g)),
    n_with_serving_ml = sum(!is.na(serving_size_ml))
  )
```

# Step 4: Per-store HEI Simulation
# Approach adapted from testing50_HEI_SCORE.R (Jus).
# Key addition: compute_hei_products() (missing from Jus) + per-store loop.

```{r step4-data}
nutrient      <- read_csv(file.path(lab_dir, "dataraw/nutrient.csv"),
                          show_col_types = FALSE)
food_nutrient_raw <- read_csv(file.path(lab_dir, "dataraw/food_nutrient.csv"),
                              col_types = cols(.default = "c"), show_col_types = FALSE)
fped <- read_xls(file.path(lab_dir, "dataraw/FPED_1720.xls"))

food_id <- unique(as.character(top50_by_category$fdc_id))

food_nutrient_joined <- food_nutrient_raw %>%
  mutate(fdc_id      = as.character(fdc_id),
         nutrient_id = as.integer(nutrient_id),
         amount      = as.numeric(amount)) %>%
  filter(fdc_id %in% food_id) %>%
  select(fdc_id, nutrient_id, amount) %>%
  left_join(nutrient, by = join_by(nutrient_id == id)) %>%
  select(fdc_id, name, amount) %>%
  group_by(name, fdc_id) %>%
  summarise(amount = mean(amount, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = name, values_from = amount)

df_hei <- top50_by_category %>%
  mutate(fdc_id = as.character(fdc_id)) %>%
  left_join(branded_food %>% select(fdc_id, ingredients), by = "fdc_id") %>%
  left_join(food_nutrient_joined, by = "fdc_id") %>%
  mutate(
    serving_g          = coalesce(serving_size_g, serving_size_ml),
    scale              = 100 / serving_g,
    calories_100g      = Energy                          * scale,
    total_fat_100g     = `Total lipid (fat)`             * scale,
    saturated_fat_100g = `Fatty acids, total saturated`  * scale,
    sodium_100g        = `Sodium, Na`                    * scale
  )
```

```{r step4-foodgroups}
contains_any <- function(text, keywords) {
  str_detect(str_to_lower(text), str_to_lower(paste(keywords, collapse = "|")))
}

df_hei <- df_hei %>%
  mutate(
    is_dairy = contains_any(ingredients, c(
      "milk", "cream", "cheese", "yogurt", "yoghurt", "whey", "lactose",
      "butter", "buttermilk", "casein", "nonfat milk", "skim milk", "whole milk"
    )),
    is_fruit = contains_any(ingredients, c(
      "apple", "orange", "grape", "lemon", "lime", "mango", "peach", "berry",
      "blueberry", "strawberry", "raspberry", "cherry", "pineapple", "banana",
      "fruit juice", "fruit puree", "juice concentrate", "cranberry", "apricot",
      "plum", "pear"
    )),
    is_vegetable = contains_any(ingredients, c(
      "tomato", "spinach", "carrot", "broccoli", "pepper", "onion", "garlic",
      "celery", "corn", "pea", "beet", "zucchini", "cucumber", "lettuce",
      "kale", "cabbage", "potato", "sweet potato", "tomatillo", "avocado", "cactus"
    )),
    is_nutseed = contains_any(ingredients, c(
      "almond", "walnut", "pecan", "cashew", "pistachio", "hazelnut", "macadamia",
      "brazil nut", "pine nut", "flax", "chia", "hemp seed", "sunflower seed",
      "pumpkin seed", "sesame seed"
    )),
    is_wholegrain = contains_any(ingredients, c(
      "whole wheat", "whole grain", "oats", "oat bran", "oat flour", "brown rice",
      "quinoa", "millet", "buckwheat", "amaranth", "barley", "rye flour",
      "whole rye", "bran", "spelt"
    )),
    is_refinedgrain = contains_any(ingredients, c(
      "enriched flour", "enriched wheat flour", "wheat flour", "bleached flour",
      "unbleached flour", "white flour", "corn starch", "rice flour", "semolina",
      "pasta", "macaroni", "noodle", "corn meal", "corn grits",
      "modified food starch", "modified corn starch"
    )),
    is_legume = contains_any(ingredients, c(
      "soy", "soybean", "tofu", "tempeh", "edamame", "lentil", "chickpea",
      "black bean", "kidney bean", "navy bean", "pinto bean", "peanut",
      "pea protein", "soy protein", "soy flour", "soy lecithin"
    )),
    is_seafood = contains_any(ingredients, c(
      "salmon", "tuna", "cod", "grouper", "tilapia", "halibut", "sardine",
      "anchovy", "shrimp", "crab", "lobster", "scallop", "clam", "oyster",
      "fish", "seafood", "seaweed", "nori", "crabmeat"
    )),
    is_meat = contains_any(ingredients, c(
      "beef", "pork", "chicken", "turkey", "lamb", "veal", "venison", "bison",
      "duck", "goose", "meat", "poultry", "prosciutto", "bacon", "sausage",
      "salami", "pepperoni", "ham"
    ))
  ) %>%
  mutate(across(starts_with("is_"), as.integer))
```

```{r step4-fped}
ingredient_fped_map <- tribble(
  ~food_group,      ~ingredient_keyword,       ~fped_desc_pattern,
  "is_dairy",       "whole milk",              "milk, whole$",
  "is_dairy",       "reduced fat milk",        "milk, reduced fat",
  "is_dairy",       "low fat milk",            "milk, low fat",
  "is_dairy",       "skim milk",               "milk, fat free|milk, skim",
  "is_dairy",       "nonfat milk",             "milk, fat free",
  "is_dairy",       "buttermilk",              "^buttermilk",
  "is_dairy",       "evaporated milk",         "milk, evaporated",
  "is_dairy",       "cream cheese",            "cream cheese$",
  "is_dairy",       "cheddar cheese",          "cheddar cheese$",
  "is_dairy",       "mozzarella",              "mozzarella cheese$",
  "is_dairy",       "parmesan",                "parmesan cheese$",
  "is_dairy",       "yogurt",                  "yogurt, NFS|yogurt, plain",
  "is_dairy",       "sour cream",              "sour cream$",
  "is_dairy",       "heavy cream",             "cream, heavy",
  "is_dairy",       "whey",                    "^whey$",
  "is_dairy",       "light cream",             "^cream, light|half and half",
  "is_dairy",       "pasteurized milk",        "^milk, NFS|^milk, whole",
  "is_fruit",       "apple juice",             "apple juice, 100%$",
  "is_fruit",       "apple",                   "^apples$|^apple, raw",
  "is_fruit",       "orange juice",            "orange juice, 100%$",
  "is_fruit",       "mango",                   "^mango, raw|^mango$",
  "is_fruit",       "mango puree",             "mango, NS|mango, raw",
  "is_fruit",       "peach",                   "^peach, raw|^peach$",
  "is_fruit",       "strawberry",              "^strawberries|strawberry, raw",
  "is_fruit",       "blueberry",               "^blueberries|blueberry, raw",
  "is_fruit",       "raspberry",               "^raspberries|raspberry, raw",
  "is_fruit",       "grape",                   "^grapes$|grape juice",
  "is_fruit",       "lemon juice",             "lemon juice$",
  "is_fruit",       "lime juice",              "lime juice$",
  "is_fruit",       "cranberry",               "^cranberries|cranberry juice",
  "is_fruit",       "pineapple",               "^pineapple, raw|pineapple juice",
  "is_fruit",       "banana",                  "^banana, raw|^banana$",
  "is_vegetable",   "tomato",                  "^tomato, raw|^tomatoes, raw",
  "is_vegetable",   "tomato paste",            "tomato paste$",
  "is_vegetable",   "tomato puree",            "tomato puree$",
  "is_vegetable",   "crushed tomatoes",        "tomatoes, crushed|tomato sauce",
  "is_vegetable",   "spinach",                 "^spinach, raw|^spinach, cooked",
  "is_vegetable",   "carrot",                  "^carrots, raw|^carrot$",
  "is_vegetable",   "broccoli",                "^broccoli, raw|^broccoli, cooked",
  "is_vegetable",   "green pepper",            "^peppers, green, raw",
  "is_vegetable",   "red pepper",              "^peppers, red, raw",
  "is_vegetable",   "jalapeno",                "^peppers, jalapeno",
  "is_vegetable",   "onion",                   "^onions, raw|^onion, raw",
  "is_vegetable",   "garlic",                  "^garlic, raw|^garlic$",
  "is_vegetable",   "celery",                  "^celery, raw|^celery$",
  "is_vegetable",   "corn",                    "^corn, yellow, cooked|^corn, canned",
  "is_vegetable",   "potato",                  "^potatoes, boiled|^potato, NS",
  "is_vegetable",   "avocado",                 "^avocado, raw|^avocado$",
  "is_vegetable",   "cucumber",                "^cucumber, raw|^cucumber$",
  "is_vegetable",   "beet",                    "^beets, cooked|^beets, raw",
  "is_vegetable",   "olives",                  "^olives|olive, ripe",
  "is_vegetable",   "green beans",             "^green beans, cooked|^snap beans",
  "is_nutseed",     "almond",                  "^almonds|almond, dry roasted",
  "is_nutseed",     "flax",                    "^flax seeds$",
  "is_nutseed",     "chia",                    "^chia seeds$",
  "is_nutseed",     "hemp",                    "^hemp seeds$",
  "is_wholegrain",  "oats",                    "^oatmeal, NS|^oat cereal",
  "is_wholegrain",  "oat bran",                "oat bran$",
  "is_wholegrain",  "whole wheat flour",       "whole wheat bread|whole wheat flour",
  "is_wholegrain",  "whole grain",             "whole grain cereal|whole grain bread",
  "is_wholegrain",  "brown rice",              "^brown rice, cooked|^brown rice$",
  "is_wholegrain",  "quinoa",                  "^quinoa, cooked|^quinoa$",
  "is_wholegrain",  "millet",                  "^millet, cooked|^millet$",
  "is_wholegrain",  "buckwheat",               "^buckwheat|buckwheat groats",
  "is_wholegrain",  "amaranth",                "^amaranth",
  "is_wholegrain",  "rye flour",               "^rye bread|rye flour",
  "is_refinedgrain","enriched wheat flour",    "white bread|enriched wheat flour",
  "is_refinedgrain","wheat flour",             "wheat flour fritter|white bread",
  "is_refinedgrain","white bread",             "^white bread$",
  "is_refinedgrain","pasta",                   "^pasta, cooked|^pasta$",
  "is_refinedgrain","macaroni",                "^macaroni, cooked|^macaroni$",
  "is_refinedgrain","corn starch",             "^corn starch|cornstarch",
  "is_refinedgrain","rice flour",              "^rice flour|rice, white, cooked",
  "is_refinedgrain","corn grits",              "^corn grits|grits, cooked",
  "is_refinedgrain","jasmine rice",            "^rice, white, cooked|^rice, NS",
  "is_refinedgrain","semolina",                "^pasta, cooked|macaroni, cooked",
  "is_legume",      "soybean",                 "^soybeans, cooked|^soybean$",
  "is_legume",      "soy milk",                "^soy milk$",
  "is_legume",      "soy protein",             "soy protein|soybean protein",
  "is_legume",      "soy lecithin",            "soy milk|^soy milk",
  "is_legume",      "pea protein",             "^peas, cooked|pea protein",
  "is_legume",      "black bean",              "^black beans, NFS|black beans, cooked",
  "is_legume",      "kidney bean",             "^kidney beans|kidney beans, cooked",
  "is_legume",      "chickpea",                "^chickpeas|garbanzo beans",
  "is_legume",      "lentil",                  "^lentils, cooked|^lentils$",
  "is_legume",      "peanut",                  "^peanuts, dry roasted|^peanuts, raw",
  "is_seafood",     "salmon",                  "^salmon, cooked|salmon, NS as to cooking",
  "is_seafood",     "tuna",                    "^tuna, NS as to cooking|^tuna, canned",
  "is_seafood",     "cod",                     "^cod, cooked|cod, NS as to cooking",
  "is_seafood",     "grouper",                 "^grouper|grouper, cooked",
  "is_seafood",     "tilapia",                 "^tilapia|tilapia, cooked",
  "is_seafood",     "shrimp",                  "^shrimp, cooked|shrimp, NS",
  "is_seafood",     "crab",                    "^crab, cooked|^crabmeat",
  "is_seafood",     "scallop",                 "^scallops|scallop, cooked",
  "is_seafood",     "halibut",                 "^halibut|halibut, cooked",
  "is_seafood",     "sardine",                 "^sardines",
  "is_meat",        "beef",                    "^beef, NS as to cut, cooked",
  "is_meat",        "pork",                    "^pork, NS as to cut, cooked",
  "is_meat",        "chicken",                 "^chicken breast, NS as to cooking",
  "is_meat",        "chicken breast",          "^chicken breast, NS as to cooking",
  "is_meat",        "turkey",                  "^turkey, nfs$",
  "is_meat",        "lamb",                    "^lamb, NS as to cut, cooked",
  "is_meat",        "bacon",                   "^bacon, NS|^bacon, cooked",
  "is_meat",        "ham",                     "^ham, NS|^ham, sliced",
  "is_meat",        "sausage",                 "^sausage, NS|^pork sausage",
  "is_meat",        "salami",                  "^salami",
  "is_meat",        "pepperoni",               "^pepperoni"
)

fped_lookup <- ingredient_fped_map %>%
  rowwise() %>%
  mutate(
    matched = list(
      fped %>%
        filter(str_detect(str_to_lower(DESCRIPTION), str_to_lower(fped_desc_pattern))) %>%
        slice(1) %>%
        select(FOODCODE, DESCRIPTION)
    )
  ) %>%
  ungroup() %>%
  unnest(matched, names_sep = "_") %>%
  rename(FOODCODE = matched_FOODCODE, fped_description = matched_DESCRIPTION) %>%
  filter(!is.na(FOODCODE)) %>%
  left_join(fped, by = "FOODCODE")

nutrient_cols <- setdiff(names(fped), c("FOODCODE", "DESCRIPTION"))

match_ingredients <- function(fdc_id_val, ingredients_val, lookup) {
  lookup %>%
    filter(str_detect(str_to_lower(ingredients_val), str_to_lower(ingredient_keyword))) %>%
    mutate(fdc_id = fdc_id_val)
}

df_with_fped <- lapply(seq_len(nrow(df_hei)), function(i) {
  match_ingredients(df_hei$fdc_id[[i]], coalesce(df_hei$ingredients[[i]], ""), fped_lookup) %>%
    mutate(gtin_upc = df_hei$gtin_upc[[i]])
}) %>% bind_rows()

df_wide <- df_with_fped %>%
  group_by(fdc_id) %>%
  summarise(
    gtin_upc            = first(gtin_upc),
    food_groups_matched = paste(unique(food_group), collapse = ", "),
    foodcodes_matched   = paste(unique(FOODCODE),   collapse = ", "),
    n_matches           = n(),
    across(all_of(nutrient_cols), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )

hei_input <- df_wide %>%
  rename_with(~ str_replace_all(., " \\(cup eq\\)| \\(oz eq\\)| \\(tsp eq\\)| \\(g\\)", "")) %>%
  left_join(
    df_hei %>% select(fdc_id, calories_100g, total_fat_100g, saturated_fat_100g, sodium_100g),
    by = "fdc_id"
  ) %>%
  mutate(
    FWHOLEFRT     = F_CITMLB + F_OTHER,
    VTOTALLEG     = V_TOTAL  + V_LEGUMES,
    VDRKGRLEG     = V_DRKGR  + V_LEGUMES,
    PFALLPROTLEG  = PF_MPS_TOTAL + PF_EGGS + PF_NUTSDS + PF_SOY + PF_LEGUMES * 4,
    PFSEAPLANTLEG = PF_SEAFD_HI  + PF_SEAFD_LOW + PF_NUTSDS + PF_SOY + PF_LEGUMES * 4,
    monopoly_g    = pmax(total_fat_100g - saturated_fat_100g, 0)
  )

cat("Products with FPED match:", n_distinct(df_with_fped$fdc_id), "of", nrow(df_hei), "\n")

n_before <- nrow(hei_input)
hei_input_valid <- hei_input %>%
  filter(
    !is.na(calories_100g), calories_100g > 0,
    !is.na(saturated_fat_100g),
    !is.na(monopoly_g)
  )
n_dropped <- n_before - nrow(hei_input_valid)
cat("Dropped", n_dropped, "of", n_before,
    "products due to missing/invalid nutrient data.\n")
```

```{r step4-functions}
# ── HEI-2020 component scoring (copied from Jus) ──────────────────────────────
score_adequacy <- function(intake, kcal, standard, max_pts) {
  density <- intake / (kcal / 1000)
  if (density == 0) return(0)
  min(max_pts * (density / standard), max_pts)
}

score_moderation_density <- function(intake, kcal, min_std, max_std, max_pts = 10) {
  density <- intake / (kcal / 1000)
  if (density <= min_std) return(max_pts)
  if (density >= max_std) return(0)
  max_pts * (1 - (density - min_std) / (max_std - min_std))
}

score_sodium <- function(sodium_mg, kcal, max_pts = 10) {
  ratio <- sodium_mg / kcal
  if (ratio <= 1.1) return(max_pts)
  if (ratio >= 2.0) return(0)
  max_pts * (1 - (ratio - 1.1) / (2.0 - 1.1))
}

score_addsug <- function(add_sugars_tsp, kcal, max_pts = 10) {
  pct <- 100 * (add_sugars_tsp * 16) / kcal
  if (pct <= 6.5) return(max_pts)
  if (pct >= 26)  return(0)
  max_pts * (1 - (pct - 6.5) / (26 - 6.5))
}

score_satfat <- function(satfat_g, kcal, max_pts = 10) {
  pct <- 100 * (satfat_g * 9) / kcal
  if (pct <= 8)  return(max_pts)
  if (pct >= 16) return(0)
  max_pts * (1 - (pct - 8) / (16 - 8))
}

score_fattyacid <- function(monopoly_g, satfat_g, max_pts = 10) {
  if (satfat_g == 0 && monopoly_g == 0) return(0)
  if (satfat_g == 0 && monopoly_g > 0)  return(max_pts)
  ratio <- monopoly_g / satfat_g
  if (ratio >= 2.5) return(max_pts)
  if (ratio <= 1.2) return(0)
  max_pts * (ratio - 1.2) / (2.5 - 1.2)
}

# ── compute_hei_products (from Jus's testing_baskets_by_store.R) ─────────────
compute_hei_products <- function(fdc_ids, grams, hei_input) {
  selected <- tibble(fdc_id = as.character(fdc_ids), grams = grams) %>%
    left_join(hei_input %>% mutate(fdc_id = as.character(fdc_id)), by = "fdc_id") %>%
    mutate(scale = grams / 100) %>%
    mutate(across(
      c(F_TOTAL, FWHOLEFRT, VTOTALLEG, VDRKGRLEG,
        G_WHOLE, G_REFINED, D_TOTAL,
        PFALLPROTLEG, PFSEAPLANTLEG,
        ADD_SUGARS, calories_100g, sodium_100g,
        saturated_fat_100g, monopoly_g),
      ~ . * scale
    ))

  totals <- selected %>%
    summarise(
      kcal          = sum(calories_100g,      na.rm = TRUE),
      F_TOTAL       = sum(F_TOTAL,            na.rm = TRUE),
      FWHOLEFRT     = sum(FWHOLEFRT,          na.rm = TRUE),
      VTOTALLEG     = sum(VTOTALLEG,          na.rm = TRUE),
      VDRKGRLEG     = sum(VDRKGRLEG,          na.rm = TRUE),
      G_WHOLE       = sum(G_WHOLE,            na.rm = TRUE),
      G_REFINED     = sum(G_REFINED,          na.rm = TRUE),
      D_TOTAL       = sum(D_TOTAL,            na.rm = TRUE),
      PFALLPROTLEG  = sum(PFALLPROTLEG,       na.rm = TRUE),
      PFSEAPLANTLEG = sum(PFSEAPLANTLEG,      na.rm = TRUE),
      ADD_SUGARS    = sum(ADD_SUGARS,         na.rm = TRUE),
      SODIUM        = sum(sodium_100g,        na.rm = TRUE),
      SATFAT        = sum(saturated_fat_100g, na.rm = TRUE),
      MONOPOLY      = sum(monopoly_g,         na.rm = TRUE)
    )

  kcal <- totals$kcal
  if (is.na(kcal) || kcal <= 0) stop("Invalid total kcal for selected products")

  scores <- c(
    fruit      = score_adequacy(totals$F_TOTAL,       kcal, 0.8,  5),
    wholefruit = score_adequacy(totals$FWHOLEFRT,     kcal, 0.4,  5),
    veg        = score_adequacy(totals$VTOTALLEG,     kcal, 1.1,  5),
    greensbean = score_adequacy(totals$VDRKGRLEG,     kcal, 0.2,  5),
    wholegrain = score_adequacy(totals$G_WHOLE,       kcal, 1.5, 10),
    dairy      = score_adequacy(totals$D_TOTAL,       kcal, 1.3, 10),
    protein    = score_adequacy(totals$PFALLPROTLEG,  kcal, 2.5,  5),
    seaplant   = score_adequacy(totals$PFSEAPLANTLEG, kcal, 0.8,  5),
    fattyacid  = score_fattyacid(totals$MONOPOLY, totals$SATFAT),
    refinedgr  = score_moderation_density(totals$G_REFINED, kcal, 1.8, 4.3),
    sodium     = score_sodium(totals$SODIUM, kcal),
    addsug     = score_addsug(totals$ADD_SUGARS, kcal),
    satfat     = score_satfat(totals$SATFAT, kcal)
  )

  list(
    total_score = sum(scores, na.rm = TRUE),
    components  = scores,
    totals      = totals
  )
}

# ── Classify products + build slot allocation (adapted from Jus) ──────────────
hei_input_classified <- hei_input_valid %>%
  distinct(fdc_id, .keep_all = TRUE) %>%
  mutate(
    primary_group = case_when(
      str_detect(food_groups_matched, "is_meat|is_seafood|is_legume|is_nutseed") ~ "protein",
      str_detect(food_groups_matched, "is_wholegrain|is_refinedgrain")           ~ "grain",
      str_detect(food_groups_matched, "is_vegetable")                            ~ "veg",
      str_detect(food_groups_matched, "is_fruit")                                ~ "fruit",
      str_detect(food_groups_matched, "is_dairy")                                ~ "dairy",
      TRUE                                                                        ~ "other"
    ),
    grain_type = case_when(
      primary_group == "grain" & str_detect(food_groups_matched, "is_wholegrain")   ~ "wholegrain",
      primary_group == "grain" & str_detect(food_groups_matched, "is_refinedgrain") ~ "refined",
      TRUE ~ NA_character_
    )
  )

# Slot allocation hardcoded by Jus: 3 veg, 2 fruit, 2 wholegrain, 1 refined, 2 protein, 2 dairy = 12 total
slot_allocation <- tibble(
  group2 = c("is_vegetable", "is_fruit", "is_wholegrain",
             "is_refinedgrain", "is_protein", "is_dairy"),
  slots  = c(3L, 2L, 2L, 1L, 2L, 2L)
)

slot_allocation

# ── Basket building + simulation (copied from Jus) ────────────────────────────
build_proportional_basket <- function(classified_input, slot_allocation,
                                      grams_per_slot = 100) {
  pool_for <- function(grp, data) {
    switch(grp,
      is_protein      = filter(data, primary_group == "protein"),
      is_vegetable    = filter(data, primary_group == "veg"),
      is_dairy        = filter(data, primary_group == "dairy"),
      is_refinedgrain = filter(data, grain_type    == "refined"),
      is_fruit        = filter(data, primary_group == "fruit"),
      is_wholegrain   = filter(data, grain_type    == "wholegrain"),
      stop(paste("Unknown group:", grp))
    )
  }

  safe_sample <- function(pool, n, label) {
    if (nrow(pool) == 0) stop(paste("Empty pool:", label))
    if (nrow(pool) < n) {
      warning(paste(label, "pool has only", nrow(pool), "products — sampling with replacement"))
      sample_n(pool, n, replace = TRUE)
    } else {
      sample_n(pool, n)
    }
  }

  basket <- map_dfr(seq_len(nrow(slot_allocation)), function(i) {
    grp  <- slot_allocation$group2[i]
    n    <- slot_allocation$slots[i]
    pool <- pool_for(grp, classified_input)
    safe_sample(pool, n, grp) %>%
      mutate(myplate_group = grp, grams = grams_per_slot)
  })

  result <- compute_hei_products(
    fdc_ids   = basket$fdc_id,
    grams     = basket$grams,
    hei_input = classified_input
  )
  result$basket     <- select(basket, fdc_id, myplate_group, grams, calories_100g)
  result$total_kcal <- result$totals$kcal
  result
}

simulate_proportional_baskets <- function(classified_input, slot_allocation,
                                          n = 100, grams_per_slot = 100,
                                          target_kcal = 2000) {
  raw <- replicate(n, {
    tryCatch(
      build_proportional_basket(classified_input, slot_allocation, grams_per_slot),
      error = function(e) NULL
    )
  }, simplify = FALSE)

  successes <- Filter(Negate(is.null), raw)
  if (length(successes) < n)
    message(n - length(successes), " baskets failed and were dropped")

  tibble(
    basket_id      = seq_along(successes),
    hei_total      = map_dbl(successes, "total_score"),
    total_kcal     = map_dbl(successes, ~ .x$totals$kcal),
    days_in_basket = total_kcal / target_kcal,
    kcal_scale     = target_kcal / total_kcal,
    basket         = map(successes, "basket")
  )
}

# ── Per-store simulation wrapper (from Jus's testing_baskets_by_store.R) ──────
pool_size_for_group <- function(grp, data) {
  switch(grp,
    is_vegetable    = sum(data$primary_group == "veg"),
    is_fruit        = sum(data$primary_group == "fruit"),
    is_wholegrain   = sum(data$grain_type    == "wholegrain", na.rm = TRUE),
    is_refinedgrain = sum(data$grain_type    == "refined",    na.rm = TRUE),
    is_protein      = sum(data$primary_group == "protein"),
    is_dairy        = sum(data$primary_group == "dairy"),
    0L
  )
}

simulate_store_baskets <- function(store_id, store_df,
                                   hei_classified = hei_input_classified,
                                   slot_alloc     = slot_allocation,
                                   n_baskets      = 50,
                                   grams_per_slot = 100,
                                   target_kcal    = 2000) {
  store_fdc_ids <- store_df %>%
    filter(!is.na(fdc_id)) %>%
    pull(fdc_id) %>%
    as.character() %>%
    unique()

  store_pool <- hei_classified %>% filter(fdc_id %in% store_fdc_ids)
  n_matched  <- nrow(store_pool)

  # Check every slot group has at least one product — empty groups cause all baskets to fail
  pool_sizes  <- sapply(slot_alloc$group2, pool_size_for_group, data = store_pool)
  empty_groups <- slot_alloc$group2[pool_sizes == 0]

  if (length(empty_groups) > 0) {
    message("Store ", store_id, ": empty group(s) [",
            paste(empty_groups, collapse = ", "), "] — skipped")
    return(NULL)
  }

  sims <- withCallingHandlers(
    simulate_proportional_baskets(
      classified_input = store_pool,
      slot_allocation  = slot_alloc,
      n                = n_baskets,
      grams_per_slot   = grams_per_slot,
      target_kcal      = target_kcal
    ),
    warning = function(w) {
      if (grepl("sampling with replacement", conditionMessage(w))) invokeRestart("muffleWarning")
    }
  )

  sims %>% mutate(store_code_uc = store_id, n_matched = n_matched)
}
```

```{r step4-per-store}
TARGET_KCAL        <- 2000
N_BASKETS_PER_STORE <- 50
GRAMS_PER_SLOT     <- 100
n_top              <- 5    # top HEI baskets to keep per store for store_top_hei

set.seed(42)

# Bad-store guard: only run stores whose CSV has an fdc_id column populated
good_stores <- names(basket_by_store)[
  map_lgl(basket_by_store, ~ "fdc_id" %in% names(.x) && any(!is.na(.x$fdc_id)))
]
cat("Stores to process:", length(good_stores), "of", length(basket_by_store), "\n")

# Run per-store simulation (returns all n_baskets per store as rows)
all_store_sims <- imap_dfr(
  basket_by_store[good_stores],
  function(store_df, store_id) {
    simulate_store_baskets(
      store_id       = store_id,
      store_df       = store_df,
      n_baskets      = N_BASKETS_PER_STORE,
      grams_per_slot = GRAMS_PER_SLOT,
      target_kcal    = TARGET_KCAL
    )
  }
)

cat("Total baskets simulated:", nrow(all_store_sims), "\n")
cat("Stores with results:", n_distinct(all_store_sims$store_code_uc), "\n")

# Per-store HEI summary (mean / median / sd)
store_channel_lookup <- store_basket_may %>%
  distinct(store_code_uc, channel_name)

store_hei_summary <- all_store_sims %>%
  group_by(store_code_uc) %>%
  summarise(
    mean_hei   = mean(hei_total,  na.rm = TRUE),
    median_hei = median(hei_total, na.rm = TRUE),
    sd_hei     = sd(hei_total,    na.rm = TRUE),
    mean_kcal  = mean(total_kcal, na.rm = TRUE),
    n_baskets  = n(),
    n_matched  = first(n_matched),
    .groups    = "drop"
  ) %>%
  left_join(store_channel_lookup, by = "store_code_uc") %>%
  arrange(desc(median_hei))

write.csv(
  store_hei_summary,
  file.path(lab_dir, "store_hei_summary.csv"),
  row.names = FALSE
)

# One price per fdc_id per store (mean across weeks if same product appears multiple times)
store_price_lookup <- store_basket_may %>%
  mutate(fdc_id = as.character(fdc_id)) %>%
  group_by(store_code_uc, fdc_id) %>%
  summarise(price = mean(price, na.rm = TRUE), .groups = "drop")

# Top-N HEI baskets per store with price
store_top_hei <- all_store_sims %>%
  group_by(store_code_uc) %>%
  arrange(desc(hei_total), .by_group = TRUE) %>%
  slice_head(n = n_top) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(
    basket_fdc_ids = map_chr(basket, ~ paste(as.character(.x$fdc_id), collapse = ",")),
    basket_price   = map2_dbl(basket, store_code_uc, function(b, sid) {
      ids    <- as.character(b$fdc_id)
      prices <- store_price_lookup %>% filter(store_code_uc == sid)
      sum(prices$price[prices$fdc_id %in% ids], na.rm = TRUE)
    })
  ) %>%
  select(store_code_uc, rank, hei_total, total_kcal, days_in_basket,
         basket_price, n_matched, basket_fdc_ids)

write.csv(
  as.data.frame(store_top_hei),
  file.path(lab_dir, "store_top_hei.csv"),
  row.names = FALSE
)

cat("Stores in store_top_hei:", n_distinct(store_top_hei$store_code_uc), "\n")

store_hei_summary %>% head(10)
```
