---
title: "Testing Code - 2026"
output:
  pdf_document: default
date: "2026-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r paths}
# Change this to your working directory
lab_dir <- "c:/Users/hughc/OneDrive/Desktop/Spring 2026/FAE Lab"

nielsen_root <- "G:/Shared drives/Nielsen Scanner/Data"
path_stores <- file.path(nielsen_root, "01_Annual_Files/stores.tsv")
path_ma_twoweeks <- file.path(nielsen_root, "MA_twoweeks")
path_branded_food <- file.path(lab_dir, "dataraw/branded_food.csv")
```

```{r libraries}
library(readr)
library(dplyr)
library(stringr)
library(tools)
```

```{r helpers}
digits_only <- function(x) {
  d <- str_remove_all(as.character(x), "[^0-9]")
  ifelse(is.na(d) | d == "", NA_character_, d)
}

pad12_last <- function(d) {
  str_sub(str_pad(d, 12, side = "left", pad = "0"), -12L)
}

add_nielsen_keys <- function(df, upc_col = "upc_chr") {
  d <- digits_only(df[[upc_col]])
  df %>%
    mutate(
      upc_digits = d,
      key_last12 = pad12_last(upc_digits),
      upc_trunc = str_sub(key_last12, 3, -1)
    )
}

# Branded Foods serving_size is usually g or ml; standardize for nutrient/HEI prep
add_serving_units <- function(df) {
  df %>%
    mutate(
      serving_size_num = as.numeric(serving_size),
      serving_unit_raw = tolower(str_trim(serving_size_unit)),
      serving_size_g = case_when(
        serving_unit_raw %in% c("g", "grm", "gram", "grams") ~ serving_size_num,
        serving_unit_raw %in% c("oz", "onz") ~ serving_size_num * 28.3495,
        TRUE ~ NA_real_
      ),
      serving_size_ml = if_else(
        serving_unit_raw %in% c("ml", "mlt"),
        serving_size_num,
        NA_real_
      ),
      serving_basis = case_when(
        serving_unit_raw %in% c("ml", "mlt") ~ "ml",
        !is.na(serving_size_g) ~ "g",
        TRUE ~ NA_character_
      )
    )
}
```

# Step 1: Nielsen two-week panel (MA)

```{r stores}
stores <- read_tsv(path_stores, show_col_types = FALSE)
```

```{r step1-load}
load_tag_and_merge <- function(folder, files = NULL, ...) {
  if (is.null(files)) {
    files <- list.files(folder, pattern = "\\.csv$", full.names = FALSE)
  }
  paths <- file.path(folder, files)
  dfs <- lapply(paths, function(p) {
    df <- read.csv(p, ...)
    df$type <- sub("_.*", "", tools::file_path_sans_ext(basename(p)))
    df
  })
  names(dfs) <- tools::file_path_sans_ext(basename(files))
  merged <- do.call(rbind, dfs)
  rownames(merged) <- NULL
  list(dfs = dfs, merged = merged)
}

result <- load_tag_and_merge(path_ma_twoweeks)
dfs <- result$dfs
merged <- result$merged %>%
  mutate(upc_chr = sprintf("%.0f", upc))

by_store <- split(merged, merged$store_code_uc)
length(by_store)
head(merged)
```

# Step 2: Link Nielsen to Branded Foods

UPC match: Nielsen `str_sub(key_last12, 3, -1)` = Branded `str_sub(gtin_upc, 2, -2)` for 12-character GTINs.

FNDDS/FPED are not joined here; `fdc_id` is retained for a future link to survey nutrients.

```{r step2-branded}
branded_food <- read_csv(path_branded_food, col_types = cols(.default = "c"))

branded_for_join <- branded_food %>%
  mutate(gtin_upc_chr = as.character(gtin_upc)) %>%
  filter(nchar(gtin_upc_chr) == 12) %>%
  mutate(gtin_upc_trunc = str_sub(gtin_upc_chr, 2, -2)) %>%
  select(
    fdc_id,
    gtin_upc = gtin_upc_chr,
    gtin_upc_trunc,
    brand_owner,
    brand_name,
    branded_food_category,
    short_description,
    serving_size,
    serving_size_unit,
    household_serving_fulltext
  )
```

```{r step2-link}
nielsen_products <- merged %>%
  group_by(upc_chr, type) %>%
  summarise(
    n_stores = n_distinct(store_code_uc),
    total_units_sold = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  add_nielsen_keys()

nielsen_fdc_linked <- nielsen_products %>%
  select(upc_chr, type, n_stores, total_units_sold, upc_trunc) %>%
  full_join(
    branded_for_join,
    by = join_by(upc_trunc == gtin_upc_trunc),
    relationship = "many-to-many"
  )

nielsen_fdc_best <- nielsen_fdc_linked %>%
  filter(!is.na(fdc_id)) %>%
  group_by(upc_chr, type) %>%
  slice(1) %>%
  ungroup()

branded_cols <- c(
  "fdc_id", "gtin_upc", "brand_owner", "brand_name",
  "branded_food_category", "short_description",
  "serving_size", "serving_size_unit", "household_serving_fulltext"
)

product_catalog <- nielsen_fdc_best %>%
  select(upc_chr, type, all_of(branded_cols)) %>%
  add_serving_units()

merged_linked <- merged %>%
  left_join(product_catalog, by = c("upc_chr", "type")) %>%
  mutate(matched_fdc = !is.na(fdc_id))

product_summary <- nielsen_products %>%
  left_join(
    nielsen_fdc_best %>% select(upc_chr, type, all_of(branded_cols)),
    by = c("upc_chr", "type")
  ) %>%
  mutate(matched = !is.na(fdc_id))

tibble(
  n_nielsen_products = nrow(nielsen_products),
  n_matched = nrow(nielsen_fdc_best),
  pct_matched = round(100 * nrow(nielsen_fdc_best) / nrow(nielsen_products), 1),
  n_merged_rows = nrow(merged_linked),
  n_merged_matched = sum(merged_linked$matched_fdc)
)
```

# Step 3: Product basket (matched, Food + Convenience)

```{r step3-channels}
channel_labels <- c(
  C = "Convenience",
  F = "Food",
  G = "Convenience with Gas Station",
  D = "Drug",
  M = "Mass Merchandiser",
  L = "Liquor"
)

stores_channels <- stores %>%
  filter(fips_state_code == 25, channel_code %in% c("F", "C")) %>%
  mutate(channel_name = recode(channel_code, !!!channel_labels)) %>%
  select(store_code_uc, channel_code, channel_name)
```

```{r step3-top50}
matched_panel <- merged_linked %>%
  filter(matched_fdc) %>%
  inner_join(stores_channels, by = "store_code_uc")

top50_by_category <- matched_panel %>%
  group_by(type, upc_chr) %>%
  summarise(
    n_stores_carrying = n_distinct(store_code_uc),
    total_units_fc = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(type) %>%
  slice_max(order_by = n_stores_carrying, n = 50, with_ties = FALSE) %>%
  ungroup() %>%
  left_join(product_catalog, by = c("upc_chr", "type"))

top50_by_category %>% count(type, name = "n_products")

write.csv(
  as.data.frame(top50_by_category),
  file.path(lab_dir, "top50_by_category.csv"),
  row.names = FALSE
)
```

```{r step3-store-basket}
store_sales_top50 <- matched_panel %>%
  inner_join(top50_by_category %>% select(type, upc_chr), by = c("type", "upc_chr")) %>%
  group_by(store_code_uc, channel_code, channel_name, type, upc_chr) %>%
  summarise(
    store_units = sum(units, na.rm = TRUE),
    .groups = "drop"
  )

top5_per_store <- store_sales_top50 %>%
  group_by(store_code_uc, channel_code, channel_name, type) %>%
  slice_max(order_by = store_units, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  select(store_code_uc, type, upc_chr)

# One row per store x product x week, carrying Nielsen's own price/units for
# that store-week directly (no cross-store or cross-week averaging).
store_basket <- matched_panel %>%
  inner_join(top5_per_store, by = c("store_code_uc", "type", "upc_chr")) %>%
  mutate(period = format(as.Date(week_end), "%b %Y")) %>%
  select(
    store_code_uc, channel_code, channel_name, type, upc_chr,
    week_end, period, units, price,
    all_of(branded_cols), serving_size_g, serving_size_ml, serving_basis
  ) %>%
  as.data.frame()

basket_by_store <- lapply(
  split(store_basket, store_basket$store_code_uc),
  as.data.frame
)

write.csv(store_basket, file.path(lab_dir, "store_basket.csv"), row.names = FALSE)

basket_dir <- file.path(lab_dir, "baskets_by_store")
dir.create(basket_dir, showWarnings = FALSE, recursive = TRUE)
for (sid in names(basket_by_store)) {
  write.csv(
    basket_by_store[[sid]],
    file.path(basket_dir, paste0(sid, ".csv")),
    row.names = FALSE
  )
}

tibble(
  n_top50_rows = nrow(top50_by_category),
  n_basket_rows = nrow(store_basket),
  n_stores = length(basket_by_store)
)
```

# Week deliverables

## TODO 0 — Matched product list (~500) for FPED work

The lab “~500 products” list is **`top50_by_category`**: matched Nielsen × Branded items, up to **50 per category** (15 categories → **747 rows**; EGGS has 47).

```{r todo0-product-list}
product_list_matched <- top50_by_category

todo0_check <- tibble(
  n_rows_product_list = nrow(product_list_matched),
  n_categories = n_distinct(product_list_matched$type),
  all_have_fdc_id = all(!is.na(product_list_matched$fdc_id)),
  all_have_gtin = all(!is.na(product_list_matched$gtin_upc))
)

todo0_check

write.csv(
  as.data.frame(product_list_matched),
  file.path(lab_dir, "product_list_matched.csv"),
  row.names = FALSE
)

# Full matched catalog (all UPC × type with a branded hit, not only top 50)
write.csv(
  as.data.frame(nielsen_fdc_best),
  file.path(lab_dir, "matched_products_all.csv"),
  row.names = FALSE
)
```

## TODO 1 — Sanity check: unmatched products

```{r todo1-unmatched}
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

unmatched_products <- product_summary %>%
  filter(!matched) %>%
  arrange(type, desc(total_units_sold))

if (file.exists(path_productdesc)) {
  product_desc <- read_tsv(path_productdesc, show_col_types = FALSE) %>%
    mutate(upc_chr = sprintf("%.0f", as.numeric(upc))) %>%
    distinct(upc_chr, .keep_all = TRUE) %>%
    select(upc_chr, product_descr)

  unmatched_products <- unmatched_products %>%
    left_join(product_desc, by = "upc_chr")
}

unmatched_by_type <- unmatched_products %>%
  group_by(type) %>%
  summarise(
    n_unmatched = n(),
    total_units_unmatched = sum(total_units_sold, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(product_summary %>% count(type, name = "n_total"), by = "type") %>%
  mutate(pct_unmatched = round(100 * n_unmatched / n_total, 1)) %>%
  arrange(desc(n_unmatched))

match_rate_by_type <- product_summary %>%
  group_by(type) %>%
  summarise(
    n_total = n(),
    n_matched = sum(matched),
    n_unmatched = sum(!matched),
    pct_matched = round(100 * mean(matched), 1),
    .groups = "drop"
  )

unmatched_by_type
match_rate_by_type

write.csv(as.data.frame(unmatched_products), file.path(lab_dir, "unmatched_products.csv"), row.names = FALSE)
write.csv(as.data.frame(unmatched_by_type), file.path(lab_dir, "unmatched_by_type.csv"), row.names = FALSE)
write.csv(as.data.frame(match_rate_by_type), file.path(lab_dir, "match_rate_by_type.csv"), row.names = FALSE)
```

## TODO 2 — Serving units (product catalog) + store-level pricing

```{r todo2-product-catalog}
write.csv(as.data.frame(product_catalog), file.path(lab_dir, "product_catalog.csv"), row.names = FALSE)

product_catalog %>%
  summarise(
    n_products = n(),
    n_with_serving_g = sum(!is.na(serving_size_g)),
    n_with_serving_ml = sum(!is.na(serving_size_ml))
  )
```

## Files to share with the lab

```{r todo-upload-list}
upload_files <- c(
  "product_list_matched.csv",
  "product_catalog.csv",
  "top50_by_category.csv",
  "unmatched_products.csv",
  "unmatched_by_type.csv",
  "match_rate_by_type.csv",
  "store_basket.csv"
)
upload_files <- file.path(lab_dir, upload_files)
data.frame(
  file = basename(upload_files),
  exists = file.exists(upload_files),
  path = upload_files
)
```
