---
title: "Testing Code - 2026"
output:
  pdf_document: default
date: "2026-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r paths}
# Change this to your working directory
lab_dir <- "C:/Users/hughc/Spring 2026/FAE Lab"

nielsen_root <- "G:/Shared drives/Nielsen Scanner/Data"
path_stores <- file.path(nielsen_root, "01_Annual_Files/stores.tsv")
path_ma_twoweeks <- file.path(nielsen_root, "MA_twoweeks")
path_branded_food <- file.path(lab_dir, "dataraw/branded_food.csv")
path_food_desc    <- file.path(lab_dir, "FoodData_Central_branded_food_csv_2026-04-30/food.csv")

week_baseline <- as.Date("2023-05-13")   # May — primary outputs
week_alt      <- as.Date("2023-09-30")   # September — parallel outputs
```

```{r libraries}
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(tools)
library(readxl)
library(purrr)
library(ggplot2)
library(scales)
```

```{r helpers}
digits_only <- function(x) {
  d <- str_remove_all(as.character(x), "[^0-9]")
  ifelse(is.na(d) | d == "", NA_character_, d)
}

pad12_last <- function(d) {
  str_sub(str_pad(d, 12, side = "left", pad = "0"), -12L)
}

add_nielsen_keys <- function(df, upc_col = "upc_chr") {
  d <- digits_only(df[[upc_col]])
  df %>%
    mutate(
      upc_digits = d,
      key_last12 = pad12_last(upc_digits),
      upc_trunc = str_sub(key_last12, 3, -1)
    )
}

# Returns total basket cost (and item count) broken out by period.
# Input: one store's basket dataframe (or any slice of store_basket).
# Each row = one product for one week, so sum(price) per period = cost of
# buying one unit of every item in the basket during that sampling week.
basket_total_price <- function(basket_df) {
  basket_df %>%
    group_by(week_end) %>%
    summarise(
      n_items     = n(),
      total_price = sum(price, na.rm = TRUE),
      .groups     = "drop"
    )
}

# Branded Foods serving_size is usually g or ml; standardize for nutrient/HEI prep
add_serving_units <- function(df) {
  df %>%
    mutate(
      serving_size_num = as.numeric(serving_size),
      serving_unit_raw = tolower(str_trim(serving_size_unit)),
      serving_size_g = case_when(
        serving_unit_raw %in% c("g", "grm", "gram", "grams") ~ serving_size_num,
        serving_unit_raw %in% c("oz", "onz") ~ serving_size_num * 28.3495,
        TRUE ~ NA_real_
      ),
      serving_size_ml = if_else(
        serving_unit_raw %in% c("ml", "mlt"),
        serving_size_num,
        NA_real_
      ),
      serving_basis = case_when(
        serving_unit_raw %in% c("ml", "mlt") ~ "ml",
        !is.na(serving_size_g) ~ "g",
        TRUE ~ NA_character_
      )
    )
}
```

# Step 1: Nielsen two-week panel (MA)

```{r stores}
stores <- read_tsv(path_stores, show_col_types = FALSE)
```

```{r step1-load}
load_tag_and_merge <- function(folder, files = NULL, ...) {
  if (is.null(files)) {
    files <- list.files(folder, pattern = "\\.csv$", full.names = FALSE)
  }
  paths <- file.path(folder, files)
  dfs <- lapply(paths, function(p) {
    df <- read.csv(p, ...)
    df$type <- sub("_.*", "", tools::file_path_sans_ext(basename(p)))
    df
  })
  names(dfs) <- tools::file_path_sans_ext(basename(files))
  merged <- do.call(rbind, dfs)
  rownames(merged) <- NULL
  list(dfs = dfs, merged = merged)
}

result <- load_tag_and_merge(path_ma_twoweeks)
dfs <- result$dfs
merged <- result$merged %>%
  mutate(upc_chr = sprintf("%.0f", upc))

by_store <- split(merged, merged$store_code_uc)
length(by_store)
head(merged)
```

# Step 2: Link Nielsen to Branded Foods

UPC match: Nielsen `str_sub(key_last12, 3, -1)` = Branded `str_sub(gtin_upc, 2, -2)` for 12-character GTINs.

FNDDS/FPED are not joined here; `fdc_id` is retained for a future link to survey nutrients.

```{r step2-branded}
branded_food <- read_csv(path_branded_food, col_types = cols(.default = "c"))

branded_for_join <- branded_food %>%
  mutate(gtin_upc_chr = as.character(gtin_upc)) %>%
  filter(nchar(gtin_upc_chr) == 12) %>%
  mutate(gtin_upc_trunc = str_sub(gtin_upc_chr, 2, -2)) %>%
  select(
    fdc_id,
    gtin_upc = gtin_upc_chr,
    gtin_upc_trunc,
    brand_owner,
    brand_name,
    branded_food_category,
    short_description,
    serving_size,
    serving_size_unit,
    household_serving_fulltext
  )
```

```{r step2-link}
nielsen_products <- merged %>%
  group_by(upc_chr, type) %>%
  summarise(
    n_stores = n_distinct(store_code_uc),
    total_units_sold = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  add_nielsen_keys()

nielsen_fdc_linked <- nielsen_products %>%
  select(upc_chr, type, n_stores, total_units_sold, upc_trunc) %>%
  full_join(
    branded_for_join,
    by = join_by(upc_trunc == gtin_upc_trunc),
    relationship = "many-to-many"
  )

nielsen_fdc_best <- nielsen_fdc_linked %>%
  filter(!is.na(fdc_id)) %>%
  group_by(upc_chr, type) %>%
  slice(1) %>%
  ungroup()

branded_cols <- c(
  "fdc_id", "gtin_upc", "brand_owner", "brand_name",
  "branded_food_category", "short_description",
  "serving_size", "serving_size_unit", "household_serving_fulltext"
)

product_catalog <- nielsen_fdc_best %>%
  select(upc_chr, type, all_of(branded_cols)) %>%
  add_serving_units()

merged_linked <- merged %>%
  left_join(product_catalog, by = c("upc_chr", "type")) %>%
  mutate(matched_fdc = !is.na(fdc_id))

product_summary <- nielsen_products %>%
  left_join(
    nielsen_fdc_best %>% select(upc_chr, type, all_of(branded_cols)),
    by = c("upc_chr", "type")
  ) %>%
  mutate(matched = !is.na(fdc_id))

tibble(
  n_nielsen_products = nrow(nielsen_products),
  n_matched = nrow(nielsen_fdc_best),
  pct_matched = round(100 * nrow(nielsen_fdc_best) / nrow(nielsen_products), 1),
  n_merged_rows = nrow(merged_linked),
  n_merged_matched = sum(merged_linked$matched_fdc)
)
```

# Step 3: Product basket (matched, Food + Convenience)

```{r step3-channels}
channel_labels <- c(
  C = "Convenience",
  F = "Food",
  G = "Convenience with Gas Station",
  D = "Drug",
  M = "Mass Merchandiser",
  L = "Liquor"
)

stores_channels <- stores %>%
  filter(fips_state_code == 25, channel_code %in% c("F", "C")) %>%
  mutate(channel_name = recode(channel_code, !!!channel_labels)) %>%
  select(store_code_uc, channel_code, channel_name)
```

```{r step3-top50}
matched_panel <- merged_linked %>%
  filter(matched_fdc) %>%
  inner_join(stores_channels, by = "store_code_uc")

top50_by_category <- matched_panel %>%
  group_by(type, upc_chr) %>%
  summarise(
    n_stores_carrying = n_distinct(store_code_uc),
    total_units_fc = sum(units, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(type) %>%
  slice_max(order_by = n_stores_carrying, n = 50, with_ties = FALSE) %>%
  ungroup() %>%
  left_join(product_catalog, by = c("upc_chr", "type"))

top50_by_category %>% count(type, name = "n_products")

write.csv(
  as.data.frame(top50_by_category),
  file.path(lab_dir, "top50_by_category.csv"),
  row.names = FALSE
)
```

```{r step3-store-basket}
store_sales_top50 <- matched_panel %>%
  inner_join(top50_by_category %>% select(type, upc_chr), by = c("type", "upc_chr")) %>%
  group_by(store_code_uc, channel_code, channel_name, type, upc_chr) %>%
  summarise(
    store_units = sum(units, na.rm = TRUE),
    .groups = "drop"
  )

top5_per_store <- store_sales_top50 %>%
  group_by(store_code_uc, channel_code, channel_name, type) %>%
  slice_max(order_by = store_units, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  select(store_code_uc, type, upc_chr)

# One row per store x product x week, carrying Nielsen's own price/units for
# that store-week directly (no cross-store or cross-week averaging).
store_basket <- matched_panel %>%
  inner_join(top5_per_store, by = c("store_code_uc", "type", "upc_chr")) %>%
  mutate(period = format(as.Date(week_end), "%b %Y")) %>%
  select(
    store_code_uc, channel_code, channel_name, type, upc_chr,
    week_end, period, units, price,
    all_of(branded_cols), serving_size_g, serving_size_ml, serving_basis
  ) %>%
  as.data.frame()

store_basket_may <- store_basket %>% filter(as.Date(week_end) == week_baseline)
store_basket_sep <- store_basket %>% filter(as.Date(week_end) == week_alt)

write_baskets <- function(sb, dir_name) {
  sb <- sb %>%
    mutate(
      price_per_100g = if_else(
        !is.na(serving_size_g) & serving_size_g > 0,
        price * 100 / serving_size_g,
        NA_real_
      )
    )
  write.csv(sb, file.path(lab_dir, paste0(dir_name, ".csv")), row.names = FALSE)
  sb_slim <- sb %>%
    select(-price, -period, -brand_name, -short_description,
           -serving_size, -serving_size_unit, -household_serving_fulltext,
           -serving_size_g, -serving_size_ml, -serving_basis)
  by_store <- lapply(split(sb_slim, sb_slim$store_code_uc), as.data.frame)
  basket_dir <- file.path(lab_dir, dir_name)
  dir.create(basket_dir, showWarnings = FALSE, recursive = TRUE)
  for (sid in names(by_store)) {
    write.csv(by_store[[sid]], file.path(basket_dir, paste0(sid, ".csv")), row.names = FALSE)
  }
  by_store
}

basket_by_store     <- write_baskets(store_basket_may, "baskets_by_store")
basket_by_store_sep <- write_baskets(store_basket_sep, "baskets_by_store_sep2023")

tibble(
  n_top50_rows    = nrow(top50_by_category),
  n_basket_may    = nrow(store_basket_may),
  n_basket_sep    = nrow(store_basket_sep),
  n_stores        = length(basket_by_store)
)
```

```{r step3-prices}
# Median price per product per period across all MA Food + Convenience stores.
# Two columns (one per sample week) so the HEI script can look up a reference
# price without needing a specific store.
product_prices <- matched_panel %>%
  inner_join(top50_by_category %>% select(type, upc_chr), by = c("type", "upc_chr")) %>%
  mutate(period_key = paste0("price_", format(as.Date(week_end), "%b%Y"))) %>%
  group_by(type, upc_chr, period_key) %>%
  summarise(price_ref = median(price, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = period_key, values_from = price_ref)

top50_by_category <- top50_by_category %>%
  left_join(product_prices, by = c("type", "upc_chr"))

write.csv(
  as.data.frame(top50_by_category),
  file.path(lab_dir, "top50_by_category.csv"),
  row.names = FALSE
)

# Quick check: how many products have a price for each period?
top50_by_category %>%
  summarise(
    n_products       = n(),
    n_price_May2023  = sum(!is.na(price_May2023)),
    n_price_Sep2023  = sum(!is.na(price_Sep2023))
  )
```

# Week deliverables

## TODO 0 — Matched product list (~500) for FPED work

The lab “~500 products” list is **`top50_by_category`**: matched Nielsen × Branded items, up to **50 per category** (15 categories → **747 rows**; EGGS has 47).

```{r todo0-product-list}
product_list_matched <- top50_by_category

todo0_check <- tibble(
  n_rows_product_list = nrow(product_list_matched),
  n_categories = n_distinct(product_list_matched$type),
  all_have_fdc_id = all(!is.na(product_list_matched$fdc_id)),
  all_have_gtin = all(!is.na(product_list_matched$gtin_upc))
)

todo0_check

write.csv(
  as.data.frame(product_list_matched),
  file.path(lab_dir, "product_list_matched.csv"),
  row.names = FALSE
)

# Full matched catalog (all UPC × type with a branded hit, not only top 50)
write.csv(
  as.data.frame(nielsen_fdc_best),
  file.path(lab_dir, "matched_products_all.csv"),
  row.names = FALSE
)
```

## TODO 1 — Sanity check: unmatched products

```{r todo1-unmatched}
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

unmatched_products <- product_summary %>%
  filter(!matched) %>%
  arrange(type, desc(total_units_sold))

if (file.exists(path_productdesc)) {
  product_desc <- read_tsv(path_productdesc, show_col_types = FALSE) %>%
    mutate(upc_chr = sprintf("%.0f", as.numeric(upc))) %>%
    distinct(upc_chr, .keep_all = TRUE) %>%
    select(upc_chr, product_descr)

  unmatched_products <- unmatched_products %>%
    left_join(product_desc, by = "upc_chr")
}

unmatched_by_type <- unmatched_products %>%
  group_by(type) %>%
  summarise(
    n_unmatched = n(),
    total_units_unmatched = sum(total_units_sold, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(product_summary %>% count(type, name = "n_total"), by = "type") %>%
  mutate(pct_unmatched = round(100 * n_unmatched / n_total, 1)) %>%
  arrange(desc(n_unmatched))

match_rate_by_type <- product_summary %>%
  group_by(type) %>%
  summarise(
    n_total = n(),
    n_matched = sum(matched),
    n_unmatched = sum(!matched),
    pct_matched = round(100 * mean(matched), 1),
    .groups = "drop"
  )

unmatched_by_type
match_rate_by_type

write.csv(as.data.frame(unmatched_products), file.path(lab_dir, "unmatched_products.csv"), row.names = FALSE)
write.csv(as.data.frame(unmatched_by_type), file.path(lab_dir, "unmatched_by_type.csv"), row.names = FALSE)
write.csv(as.data.frame(match_rate_by_type), file.path(lab_dir, "match_rate_by_type.csv"), row.names = FALSE)
```

## TODO 3 — Match sanity check: Nielsen vs Branded Foods descriptions

```{r todo3-match-sanity}
path_productdesc <- file.path(nielsen_root, "01_Annual_Files/productdesc.tsv")

if (file.exists(path_productdesc)) {
  product_desc_all <- read_tsv(path_productdesc, show_col_types = FALSE) %>%
    mutate(upc_chr = sprintf("%.0f", as.numeric(upc))) %>%
    distinct(upc_chr, .keep_all = TRUE) %>%
    select(upc_chr, nielsen_desc = product_descr)

  match_check <- top50_by_category %>%
    left_join(product_desc_all, by = "upc_chr") %>%
    mutate(
      branded_label = paste(
        coalesce(brand_owner, ""),
        coalesce(brand_name, ""),
        coalesce(branded_food_category, ""),
        sep = " | "
      )
    ) %>%
    select(
      type,
      upc_chr,
      gtin_upc,
      nielsen_desc,
      brand_owner,
      brand_name,
      branded_label,
      branded_category = branded_food_category
    ) %>%
    arrange(type, upc_chr)

  print(match_check, n = Inf)

  write.csv(
    as.data.frame(match_check),
    file.path(lab_dir, "match_sanity_check.csv"),
    row.names = FALSE
  )
} else {
  message("productdesc.tsv not found — skipping match sanity check")
}
```

## TODO 2 — Serving units (product catalog) + store-level pricing


```{r todo2-product-catalog}
write.csv(as.data.frame(product_catalog), file.path(lab_dir, "product_catalog.csv"), row.names = FALSE)

product_catalog %>%
  summarise(
    n_products = n(),
    n_with_serving_g = sum(!is.na(serving_size_g)),
    n_with_serving_ml = sum(!is.na(serving_size_ml))
  )
```


## JUS CODE
```{r run-hei-pipeline, warning = FALSE}
source(file.path(lab_dir, "testing_basket_by_store_v2.R"))

# all_store_sims has what we want
```

## Basket Cost

```{r basket-cost}
all_store_sims <- all_store_sims %>%
  mutate(
    basket_cost      = map_dbl(basket, function(b) {
      sum(b$price_per_100g * b$grams / 100, na.rm = TRUE)
    }),
    # Cost to obtain 2000 kcal from this basket (normalizes for caloric density)
    basket_cost_norm = if_else(total_kcal > 0, basket_cost / total_kcal * 2000, NA_real_)
  )

# Healthy baskets only (HEI > 80) — used for all downstream cost analysis
healthy_baskets <- all_store_sims %>% filter(hei_total > 80)

all_store_sims %>%
  summarise(
    n_baskets        = n(),
    n_healthy        = sum(hei_total > 80, na.rm = TRUE),
    pct_healthy      = round(100 * mean(hei_total > 80, na.rm = TRUE), 1),
    mean_cost        = round(mean(basket_cost, na.rm = TRUE), 2),
    mean_cost_norm   = round(mean(basket_cost_norm, na.rm = TRUE), 2),
    median_cost      = round(median(basket_cost, na.rm = TRUE), 2),
    sd_cost          = round(sd(basket_cost, na.rm = TRUE), 2)
  )
```

## Graphs

```{r graphs-cost-distribution}
ggplot(all_store_sims, aes(x = basket_cost_norm)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Distribution of Simulated Basket Costs (per 2000 kcal)",
    x     = "Basket Cost per 2000 kcal ($)",
    y     = "Count"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r graphs-cost-vs-hei}
ggplot(all_store_sims, aes(x = basket_cost_norm, y = hei_total)) +
  geom_point(alpha = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "firebrick") +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Basket Cost vs. HEI-2020 Score",
    x     = "Basket Cost per 2000 kcal ($)",
    y     = "HEI-2020 Total Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r graphs-store-cost-hei}
store_cost_summary <- healthy_baskets %>%
  group_by(store_code, channel_name) %>%
  summarise(
    mean_hei       = mean(hei_total, na.rm = TRUE),
    mean_cost      = mean(basket_cost, na.rm = TRUE),
    mean_cost_norm = mean(basket_cost_norm, na.rm = TRUE),
    .groups        = "drop"
  ) %>%
  mutate(store_code = as.character(store_code)) %>%
  left_join(
    stores %>%
      select(store_code_uc, parent_code) %>%
      mutate(store_code_uc = as.character(store_code_uc)),
    by = c("store_code" = "store_code_uc")
  ) %>%
  mutate(parent_code = as.character(parent_code))

# Group low-frequency chains into "Other" for a readable legend
top_parents <- store_cost_summary %>%
  count(parent_code, sort = TRUE) %>%
  slice_head(n = 8) %>%
  pull(parent_code)

store_cost_summary <- store_cost_summary %>%
  mutate(
    chain_group = if_else(parent_code %in% top_parents, parent_code, "Other"),
    chain_group = factor(chain_group, levels = c(top_parents, "Other"))
  )

ggplot(store_cost_summary, aes(x = mean_cost_norm, y = mean_hei, color = chain_group)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "firebrick", linewidth = 0.8) +
  scale_color_viridis_d(option = "mako", name = "Parent Code") +
  scale_x_continuous(labels = dollar) +
  labs(
    title    = "Store-Level Mean Basket Cost vs. Mean HEI-2020 Score",
    subtitle = "Healthy baskets only (HEI > 80); cost normalized to 2000 kcal",
    x        = "Mean Basket Cost per 2000 kcal ($)",
    y        = "Mean HEI-2020 Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```

# Extended Graphs

## Geographic and county setup

```{r geo-setup}
library(usmap)
library(viridis)
library(scales)
library(patchwork)

# MA county FIPS → 5-digit code, label, and 2022 ACS median household income
# Source: U.S. Census Bureau, ACS 5-Year Estimates 2022, Table B19013
ma_county_ref <- tibble(
  fips_county_code = c(1L, 3L, 5L, 7L, 9L, 11L, 13L, 15L, 17L, 19L, 21L, 23L, 25L, 27L),
  county_label     = c("Barnstable","Berkshire","Bristol","Dukes","Essex",
                       "Franklin","Hampden","Hampshire","Middlesex","Nantucket",
                       "Norfolk","Plymouth","Suffolk","Worcester"),
  median_hhincome  = c(79811, 57786, 68621, 87050, 83627,
                       61275, 56925, 65361, 104658, 108263,
                       101174, 86616, 73000, 73788)
) %>%
  mutate(fips = paste0("25", str_pad(fips_county_code, 3, pad = "0")))

# Add county info to each panel store
stores_geo <- stores %>%
  filter(fips_state_code == 25, channel_code %in% c("F", "C")) %>%
  mutate(
    channel_name     = recode(channel_code, !!!channel_labels),
    fips_county_code = as.integer(fips_county_code)
  ) %>%
  select(store_code_uc, channel_code, channel_name, fips_county_code) %>%
  left_join(ma_county_ref, by = "fips_county_code")

# Join county info onto per-store cost summary (healthy baskets only)
store_cost_geo <- store_cost_summary %>%
  left_join(
    stores_geo %>%
      mutate(store_code_uc = as.character(store_code_uc)) %>%
      select(store_code_uc, fips, fips_county_code, county_label, median_hhincome),
    by = c("store_code" = "store_code_uc")
  )

# County-level aggregation: mean, 25th and 75th percentile of store costs (normalized)
county_cost <- store_cost_geo %>%
  group_by(fips, county_label, median_hhincome) %>%
  summarise(
    n_stores       = n(),
    mean_cost_norm = mean(mean_cost_norm, na.rm = TRUE),
    q25_cost_norm  = quantile(mean_cost_norm, 0.25, na.rm = TRUE),
    q75_cost_norm  = quantile(mean_cost_norm, 0.75, na.rm = TRUE),
    .groups        = "drop"
  )
```

## Price: Basket cost by county (boxplot)

```{r graph-cost-by-county}
store_cost_geo %>%
  filter(!is.na(county_label)) %>%
  mutate(county_label = reorder(county_label, mean_cost_norm, FUN = median)) %>%
  ggplot(aes(x = county_label, y = mean_cost_norm, fill = county_label)) +
  geom_boxplot(alpha = 0.8, outlier.size = 0.8, show.legend = FALSE) +
  scale_fill_viridis_d(option = "mako") +
  coord_flip() +
  scale_y_continuous(labels = dollar) +
  labs(
    title    = "Basket Cost by MA County",
    subtitle = "Healthy baskets only (HEI > 80); cost per 2000 kcal",
    x        = NULL,
    y        = "Mean Basket Cost per 2000 kcal ($)"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title    = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```

## Price: Map of mean basket cost by county

```{r graph-cost-map}
plot_usmap("counties", include = "MA",
           data = county_cost %>% filter(!is.na(mean_cost_norm)),
           values = "mean_cost_norm") +
  scale_fill_viridis_c(
    option   = "mako",
    begin    = 0.2, end = 0.9,
    na.value = "grey90",
    labels   = dollar,
    name     = "Mean Basket\nCost ($)"
  ) +
  labs(
    title    = "Average Cost of Healthy Basket Across Massachusetts Counties",
    subtitle = "Per 2000 kcal; healthy baskets only (HEI > 80)"
  ) +
  theme(
    legend.position  = "right",
    panel.background = element_rect(color = "black", fill = "white"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle    = element_text(hjust = 0.5, size = 11)
  )
```

## Price: County cost distribution — mean, 25th, 75th percentile maps

```{r graph-cost-map-quantiles}
make_cost_map <- function(fill_col, title) {
  plot_usmap("counties", include = "MA",
             data = county_cost %>% filter(!is.na(.data[[fill_col]])),
             values = fill_col) +
    scale_fill_viridis_c(
      option   = "mako",
      begin    = 0.2, end = 0.9,
      na.value = "grey90",
      labels   = dollar,
      name     = "Cost ($)"
    ) +
    labs(title = title) +
    theme(
      legend.position  = "right",
      panel.background = element_rect(color = "black", fill = "white"),
      plot.title       = element_text(hjust = 0.5, face = "bold", size = 11)
    )
}

p_q25  <- make_cost_map("q25_cost_norm",  "25th Percentile")
p_mean <- make_cost_map("mean_cost_norm", "Mean")
p_q75  <- make_cost_map("q75_cost_norm",  "75th Percentile")

(p_q25 | p_mean | p_q75) +
  plot_annotation(
    title    = "Basket Cost by MA County — Spread Across Stores",
    subtitle = "Healthy baskets only (HEI > 80); Massachusetts NielsenIQ Scanner Data",
    theme    = theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 14),
      plot.subtitle = element_text(hjust = 0.5, size = 11)
    )
  )
```

## Price: County mean basket cost with 95% CI

```{r graph-cost-ci}
# Compute per-county mean, SE, and 95% CI from store-level normalized costs
county_ci <- store_cost_geo %>%
  filter(!is.na(county_label), !is.na(mean_cost_norm)) %>%
  group_by(county_label) %>%
  summarise(
    n        = n(),
    mean_c   = mean(mean_cost_norm, na.rm = TRUE),
    se       = sd(mean_cost_norm,   na.rm = TRUE) / sqrt(n()),
    lower    = mean_c - 1.96 * se,
    upper    = mean_c + 1.96 * se,
    .groups  = "drop"
  ) %>%
  mutate(county_label = reorder(county_label, mean_c))

statewide_mean <- mean(county_ci$mean_c)

ggplot(county_ci, aes(x = county_label, y = mean_c,
                      ymin = lower, ymax = upper,
                      color = mean_c > statewide_mean)) +
  geom_pointrange(size = 0.6, linewidth = 0.7) +
  geom_hline(yintercept = statewide_mean, linetype = "dashed",
             color = "firebrick", linewidth = 0.8) +
  annotate("text", x = 0.6, y = statewide_mean,
           label = paste0("State mean: $", round(statewide_mean, 2)),
           hjust = 0, vjust = -0.5, size = 3.5, color = "firebrick") +
  scale_color_manual(values = c("TRUE" = "#1f77b4", "FALSE" = "#2ca02c"),
                     labels = c("TRUE" = "Above state mean", "FALSE" = "Below state mean"),
                     name   = NULL) +
  scale_y_continuous(labels = dollar) +
  coord_flip() +
  labs(
    title    = "Mean Basket Cost by MA County (95% CI)",
    subtitle = "Healthy baskets only (HEI > 80); cost per 2000 kcal",
    x        = NULL,
    y        = "Mean Basket Cost per 2000 kcal ($)"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title    = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "bottom")
```

## Price: Product price vs. availability (scatter by category)

```{r graph-price-availability}
top50_by_category %>%
  filter(!is.na(price_May2023)) %>%
  ggplot(aes(x = n_stores_carrying, y = price_May2023, color = type)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_viridis_d(option = "mako") +
  scale_y_continuous(labels = dollar) +
  labs(
    title  = "Product Price vs. Availability Across MA Food/Convenience Stores",
    x      = "Number of Stores Carrying Product",
    y      = "Median Price — May 2023 ($)",
    color  = "Category"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title      = element_text(hjust = 0.5, face = "bold"),
        legend.position = "right")
```

## Stores and products: Category availability (bar chart)

```{r graph-category-bar}
top50_by_category %>%
  group_by(type) %>%
  summarise(mean_n_stores = mean(n_stores_carrying, na.rm = TRUE), .groups = "drop") %>%
  mutate(type = reorder(type, mean_n_stores)) %>%
  ggplot(aes(x = type, y = mean_n_stores, fill = type)) +
  geom_col(show.legend = FALSE) +
  scale_fill_viridis_d(option = "mako") +
  coord_flip() +
  labs(
    title = "Mean Stores Carrying Each Product Category (Top-50 products)",
    x     = NULL,
    y     = "Mean Number of Stores"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Stores and products: Product × store heatmap (selected category)

```{r graph-heatmap}
HEATMAP_CATEGORY <- "BREAD"

heatmap_products <- top50_by_category %>%
  filter(type == HEATMAP_CATEGORY) %>%
  slice_max(total_units_fc, n = 15, with_ties = FALSE) %>%
  mutate(product_label = coalesce(short_description, brand_name, upc_chr)) %>%
  select(upc_chr, product_label)

heatmap_data <- store_sales_top50 %>%
  filter(type == HEATMAP_CATEGORY) %>%
  semi_join(heatmap_products, by = "upc_chr") %>%
  left_join(heatmap_products, by = "upc_chr") %>%
  mutate(
    store_code_uc = as.character(store_code_uc),
    log_units     = log1p(store_units)
  )

# Order stores by total log-units so the most active stores cluster left
store_order <- heatmap_data %>%
  group_by(store_code_uc) %>%
  summarise(total = sum(log_units), .groups = "drop") %>%
  arrange(desc(total)) %>%
  pull(store_code_uc)

heatmap_data %>%
  mutate(store_code_uc = factor(store_code_uc, levels = store_order)) %>%
  ggplot(aes(x = store_code_uc, y = product_label, fill = log_units)) +
  geom_tile(color = NA) +
  scale_fill_viridis_c(option = "mako", begin = 0.1, end = 0.9,
                       name = "log(Units + 1)", na.value = "grey90") +
  scale_x_discrete(labels = NULL) +
  labs(
    title    = paste("Product × Store Heatmap —", HEATMAP_CATEGORY),
    subtitle = "Top 15 products by total units; each column = one store (ordered by activity)",
    x        = "Store →",
    y        = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title   = element_text(hjust = 0.5, face = "bold"),
        axis.text.x  = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid   = element_blank())
```

## Stores and products: Map of Food vs. Convenience stores by county

```{r graph-store-channel-map}
county_channel <- stores_geo %>%
  filter(!is.na(fips)) %>%
  group_by(fips, county_label, channel_name) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(fips, county_label) %>%
  mutate(n_county = sum(n), pct = n / sum(n)) %>%
  ungroup()

pct_food <- county_channel %>%
  filter(channel_name == "Food") %>%
  select(fips, county_label, n_food = n, pct_food = pct, n_county)

plot_usmap("counties", include = "MA",
           data = pct_food,
           values = "pct_food") +
  scale_fill_viridis_c(
    option   = "mako",
    begin    = 0.2, end = 0.9,
    na.value = "grey90",
    labels   = percent_format(accuracy = 1),
    name     = "% Food\nStores"
  ) +
  labs(
    title    = "Share of Food vs. Convenience Stores by MA County",
    subtitle = "Massachusetts, NielsenIQ Scanner Data"
  ) +
  theme(
    legend.position  = "right",
    panel.background = element_rect(color = "black", fill = "white"),
    plot.title       = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle    = element_text(hjust = 0.5, size = 11)
  )
```

## Price: Basket cost vs. county median household income

```{r graph-cost-vs-income}
county_cost %>%
  filter(!is.na(median_hhincome), !is.na(mean_cost_norm)) %>%
  ggplot(aes(x = median_hhincome, y = mean_cost_norm)) +
  # geom_point(aes(size = n_stores), alpha = 0.7, color = "#1b7837") +
  geom_point(alpha = 0.7, color = "#1b7837") +
  geom_smooth(method = "lm", se = TRUE, color = "firebrick", linewidth = 0.8) +
  geom_text(aes(label = county_label), vjust = -0.8, size = 3,
            check_overlap = TRUE) +
  scale_x_continuous(labels = dollar) +
  scale_y_continuous(labels = dollar) +
  scale_size_continuous(name = "# Stores", range = c(2, 8)) +
  labs(
    title    = "Basket Cost vs. Median Household Income by MA County",
    subtitle = "Cost per 2000 kcal; healthy baskets (HEI > 80); income: 2022 ACS 5-year estimates",
    x        = "Median Household Income ($)",
    y        = "Mean Basket Cost per 2000 kcal ($)"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title    = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```
